---
title: "Psammoneis HGT project"
author: "Cory Gargas"
date: "11/2/2021"
format:
  html:
    code-fold: true
    code-tools: true
    code-link: true
    theme: 
      light: flatly
      dark: darkly
---

```{r}
#| label: setup
#| include: false
knitr::opts_chunk$set(echo = TRUE)
pkgs <- c(
  "ape",
  "forcats",
  "aplot",
  "grafify",
  "report",
  "skimr",
  "janitor",
  "glue",
  "ggridges",
  "ggdist",
  "ggtext",
  "ggside",
  "ggtree",
  "GenomicRanges",
  "here",
  "reshape2",
  "tidyverse",
  "patchwork",
  "seqinr",
  "coRdon",
  "data.table",
  "TidyDensity",
  "arrow",
  "RSQLite"
) # Names of packages we want to load
pkgs_tidylog <- c(
  "report",
  "skimr",
  "janitor",
  "glue",
  "ggridges",
  "ggdist",
  "ggtext",
  "ggside",
  "ggtree",
  "GenomicRanges",
  "here",
  "reshape2",
  "tidyverse",
  "patchwork",
  "seqinr",
  "coRdon",
  "data.table",
  "TidyDensity",
  "arrow",
  "RSQLite",
  "tidylog"
) # same as above, but with tidylog if we want more verbose output when using tidy.

pacman::p_load(
  char = pkgs,
  install = TRUE,
  character.only = TRUE
) # Load our pacakges

# Load pkgs, plus tidylog
# pacman::p_load(char = pkgs_tidylog, 
#                install = TRUE, 
#                character.only = TRUE)
#                
rm(
  pkgs,
  pkgs_tidylog
)

count_nonNA_func <- function(x) sum(!is.na(x)) 
count_NA_func <- function(x) sum(is.na(x)) 

```

Note: the full code for all scripts mentioned in this document can be found in the appendix at the end of this RMD file.

# Intergenic region - HGT index evaluation

<hr />

Starting off, we downloaded our reference sequences (sseqid) which we will blast our intergenic sequences (qseqid) against. We downloaded the reviewed Swiss-Prot sequences for Eukaryote, Archaea, and Bacteria protein sequences from UniProt as fasta files.\
We then concatenated the fasta files for Archaea and bacteria into a single fasta file.

Below, we describe the code used on the AHPCC computing cluster to conduct this analysis (See 'submit_intergenic_HGT_DMNDblast_uniprot_workflow_psam_768gb.slurm' for the actual slurm file).

## AHPCC analysis walkthrough

First, we extract the CDS annotated regions of each .gff file from our bacterial genomes that were sequenced from our Psammoneis culture. We performed this using the Rscript 'gff_filter_rename_bac.R'.

We then use bedtools getfasta to extract the fasta sequences for those CDS annotations:

    bedtools getfasta -fi 000000F.quiver_pilon.n_f.fasta -bed 0F.gff -fo 000000F.fa -s
    bedtools getfasta -fi 000001F.quiver_pilon.n_f.fasta -bed 1F.gff -fo 000001F.fa -s
    bedtools getfasta -fi 000002F.quiver_pilon.n_f.fasta -bed 2F.gff -fo 000002F.fa -s
    bedtools getfasta -fi 000003F.quiver_pilon.n_f.fasta -bed 3F.gff -fo 000003F.fa -s

Then we translate those bacterial sequences into amino acid (AA) sequences using the Rscript 'translate_cds.R':

    Rscript translate_cds.R

Next, we need to get the intergenic regions of our Psammoneis genome. We start off by generating our genomefile for bedtools. Using Samtools faidx we generate a .fai file, then use cut to extract the 1st two columns from the .fai file (e.g., chromosome name, number of bases per chromosome). Lastly, we sort it to match the sorted .gff file we generate later.

        samtools faidx psam.final_assembly.Quiver_Pilon_Reapr_GapCloser.fa 

        cut -f1,2 psam.final_assembly.Quiver_Pilon_Reapr_GapCloser.fa.fai > sizes.genome

        sort -k 3 sizes.genome > sizes_sorted.genome

alternatively, you can use the following instead of the samtools method to generate a genome file:

        grep contig result_sorted.gff | cut -f 1,5 > sizes_sorted.genome

then we need to remove the fasta sequences from the end of our genome

This gives us the line number that the fasta sequences starts at.

        egrep -n -m 1 '##FASTA' psamm_final.round4.all.gff

Then we use head with the value we got, minus one, (e.g., line_number - 1) to just give us the gff lines

        head -n 1070595 psamm_final.round4.all.gff > result.gff

Next we need to sort our .gff file so bedtools doesn't get mad at us.

        bedtools sort -i result.gff > result_sorted.gff

and remove contig annotations from our .gff (or else we won't get an complement output)

        grep -v contig result_sorted.gff > result_sorted_final.gff

next, we run bedtools complement to get our intergenic .gff file

        bedtools complement -i result_sorted_final.gff -g sizes_sorted.genome > psamm_complement.txt

Finally, we run bedtools getfasta on our genome, using the complement .gff file to extract sequences for our intergenic regions.

        bedtools getfasta -fi psam.final_assembly.Quiver_Pilon_Reapr_GapCloser.fa -bed psamm_complement.txt -fo psamm_complement.fa

Next, we run orfipy to get our ORFs (open reading frames) from our intergenic regions.

        orfipy psamm_complement_intergenic.fa --pep psamm_complement_ORFs_orfipy.faa --dna psamm_complement_ORFs_orfipy.fa  --strand b --between-stops --include-stop --min 30

We also make delimited files of the headers for both .faa and .fa ORF files for downstream analysis in R.

    grep '>' psamm_complement_ORFs_orfipy.faa > psamm_complement_ORFs_orfipy.faa.headers.delim
    grep '>' psamm_complement_ORFs_orfipy.fa > psamm_complement_ORFs_orfipy.fa.headers.delim

Make some directories for our intergenic files

        mkdir 03_psam_intergenic

        mv psamm_complement_edit.bed result_sorted_final.gff psamm.genomeFile psamm_complement_intergenic.fa orfipy_psamm_complement.fa_out 03_psam_intergenic/

Now that we have our ORFs from the intergenic regions of our Psammoneis genome, we need to extract CDS sequences from both Psammoneis and the bacterial genomes associated with it. We do this using the Rscript 'gff_filter_rename_bac.R' and bedtools getfasta:

    Rscript gff_filter_rename_bac.R

    bedtools getfasta -fi 000000F.quiver_pilon.n_f.fasta -bed 0F.gff -fo 000000F.fa -s
    bedtools getfasta -fi 000001F.quiver_pilon.n_f.fasta -bed 1F.gff -fo 000001F.fa -s
    bedtools getfasta -fi 000002F.quiver_pilon.n_f.fasta -bed 2F.gff -fo 000002F.fa -s
    bedtools getfasta -fi 000003F.quiver_pilon.n_f.fasta -bed 3F.gff -fo 000003F.fa -s

Next we run our 'translate_cds.R' on the fasta files from above and make new directrories for our output files:

        Rscript translate_cds.R

        #make output directories for output
        mkdir 01_genomes_CDS
        mkdir 02_genomes_CDS_translated 

        # copy all output files from R
        mv *_CDS.fa 01_genomes_CDS/
        mv  *.faa 02_genomes_CDS_translated/

Then we append our bacterial CDS AA sequences to our .fna file for blast input

        cat 02_genomes_CDS_translated/00000*F_gff.faa >> uniprot_reviewed_Arc_2157_Bac_2.fasta

Next up we make our Diamond blast databases using the UniProt protein sequences we downloaded at the beginning:

    echo 'Make Diamond DB using NR + 0-3F bacterial genomes'
    diamond makedb --in uniprot_reviewed_Arc_2157_Bac_2.fasta -d uniprot_prok
    diamond makedb --in uniprot-reviewed_yes+taxonomy_2759_eukaryota.fasta -d uniprot_euk

    mkdir uniprot_dmnd/

    mv *.dmnd uniprot_dmnd/

Now we can finally run Diamond blastp to get our donor and recipient blast outputs:

    diamond blastp -f 6 -q orfipy_psamm_complement_intergenic.fa_out/psamm_complement_ORFs_orfipy.faa  -d uniprot_dmnd/uniprot_prok.dmnd -k 5 -o intergenic_dmnd_uniprot_prok.out.tsv -v -p 4 -b12 -c1

    diamond blastp -f 6 -q orfipy_psamm_complement_intergenic.fa_out/psamm_complement_ORFs_orfipy.faa  -d uniprot_dmnd/uniprot_euk.dmnd -k 5 -o intergenic_dmnd_uniprot_euk.out.tsv -v -p 4 -b12 -c1

    mkdir dmnd_blastp_hgt_output
    cp *.out.tsv dmnd_blastp_hgt_output

    mkdir DMNDblastp_uniprot_output_psam/

    # copy the output files to directory
    cp -r result.gff result_sorted.gff psamm.genomeFile psamm_complement.txt orfipy_psamm_complement_intergenic.fa_out psamm_complement_edit.bed uniprot_dmnd/ dmnd_blastp_hgt_output/ 02_genomes_CDS_translated/ 01_genomes_CDS/ 03_psam_intergenic/psam_intergenic_and_CDS.faa DMNDblastp_uniprot_output_psam/

## HGT index output and analysis

Next we use the R code below to process our output files and calculate the HGT index for our hits:

```{r}
#| label: Intergenic HGT index 
#| include: false


colnames_donor <- c(
  "qseqid",
  "sseqid_d",
  "pident_d",
  "length_d",
  "mismatch_d",
  "gapopen_d",
  "qstart_d",
  "qend_d",
  "sstart_d",
  "send_d",
  "evalue_d",
  "bitscore_d"
)
colnames_recipients <- c(
  "qseqid",
  "sseqid_r",
  "pident_r",
  "length_r",
  "mismatch_r",
  "gapopen_r",
  "qstart_r",
  "qend_r",
  "sstart_r",
  "send_r",
  "evalue_r",
  "bitscore_r"
)

ORF_headers <-
  read_delim(
    file = here(
      "data",
      "01_psam_intergenic_HGT_index",
      "psam_genome",
      "orfipy_psamm_complement_intergenic.fa_out",
      "psamm_complement_ORFs_orfipy_headers.delim"
    ),
    delim = " ",
    col_names = c("qseqid", "coords_strand", "type", "length", "frame", "start", "stop")
  ) %>%
  mutate(across(
    .cols = everything(),
    .fns = ~ str_remove(
      string = .x,
      pattern = "^>|^type:|^length:|^frame:|^start:|^stop:"
    )
  )) %>%
  separate(
    qseqid,
    sep = ":",
    into = c("chrm", "length_orf"),
    remove = F
  ) %>% # split and clean qseqid
  separate(
    length_orf,
    sep = "_",
    into = c("length_orf", "orf_id"),
    remove = T
  ) %>% # split ORF names 2
  mutate(length_orf = str_replace_all(
    string = length_orf,
    pattern = "[/(/)$]",
    replacement = ""
  )) %>% # remove () from orf length
  mutate(chrm = str_replace_all(
    string = chrm,
    pattern = "_quiver_pilon$",
    replacement = ""
  )) %>% # remove quiver pilon from chrm
  mutate(
    strand = str_extract_all(
      coords_strand,
      "\\(\\+\\)|\\(\\-\\)"
    )
  ) %>%
  mutate(
    strand = str_replace_all(
      string = strand,
      pattern = "^\\(|\\)$",
      replacement = ""
    )
  ) %>% # split coords_strand
  mutate(
    coords_strand = str_replace_all(
      string = coords_strand,
      pattern = "^\\[|\\]|\\(|\\+\\)$|\\-\\)$",
      replacement = ""
    )
  ) %>% # split coords_strand
  dplyr::rename(
    coords = coords_strand
  ) 

uniprot_reviewed <-
  read_tsv(
    file = here(
      "data",
      "uniprot-reviewed_yes.tsv"
    ),
    col_names = T
  ) %>%
  as.data.frame() %>%
  clean_names()


# get the directory our bacterial genome GFF annotation files are in
bac_gff_directory <-
  here(
    "data",
    "01_psam_intergenic_HGT_index",
    "psam_bac_genomes"
  )


# list the gff files in each directory as a variable
bac_gff_files <-
  list.files(
    path = bac_gff_directory,
    full.names = T,
    pattern = "\\.gff"
  )


# Here we use fread, lapply, and bind_rows to read in all of our GFF files and bind them by row (i.e., concatenating them) into a single data frame.
bac_gff <-
  bind_rows(lapply(
    bac_gff_files,
    fread,
    skip=1,
    header=F,
    col.names = c(
      "chrm_bac",
      "source",
      "feature_type",
      "feature_start",
      "feature_end",
      "score",
      "strand",
      "phase",
      "attributes"
    )
  )) %>%
  filter(
    feature_type == "CDS"
  ) %>% 
  mutate(
    attributes = str_remove(
      attributes,
      "ID\\=fig\\|203682\\.28\\.peg\\.|ID\\=fig\\|41295\\.5\\.peg\\.|ID\\=fig\\|89373\\.5\\.peg\\.|ID\\=fig\\|1813606\\.4\\.peg\\."
    )
  ) %>% 
  separate(
    col = attributes,
    into = c("bac_gen_peg", "bac_gen_gene_annotation"),
    sep = ";",
    extra = "drop"
  ) %>% 
  mutate(
    bac_gen_gene_annotation = str_remove(
      bac_gen_gene_annotation,
      "Name\\="
    )
  ) %>% 
  mutate(
    length_bac = paste(
      feature_start,
      "-",
      feature_end,
      "(",
      strand,
      ")",
      sep = ""
    )
  ) %>%
  select(-c(
    "feature_type",
    "feature_start",
    "feature_end",
    "score",
    "strand"
  )) %>% 
  mutate(
    chrm_bac = str_remove(
      chrm_bac,
      pattern = "^00000"
    )
  ) 

# import our eukaryotic HGT blast output
euk <-
  read_tsv(
    file = here(
      "data",
      "01_psam_intergenic_HGT_index",
      "DMNDblastp_uniprot_output_psam",
      "dmnd_blastp_hgt_output",
      "intergenic_dmnd_uniprot_euk.out.tsv"
    ),
    col_names = colnames_recipients,
    progress = T
  ) %>%
  group_by(qseqid) %>% # group our data by sqeqid
  top_n(1, bitscore_r) %>% # select the top blast search result by bitscore for each seq_id
  dplyr::slice(1) %>%
  ungroup()

# import our prokaryotic HGT blast output
prok <-
  read_tsv(
    file = here(
      "data",
      "01_psam_intergenic_HGT_index",
      "DMNDblastp_uniprot_output_psam",
      "dmnd_blastp_hgt_output",
      "intergenic_dmnd_uniprot_prok.out.tsv"
    ),
    col_names = colnames_donor,
    progress = T
  ) %>%
  group_by(qseqid) %>% # group our data by sqeqid
  top_n(1, bitscore_d) %>% # select the top blast search result by bitscore for each seq_id
  dplyr::slice(1) %>%
  ungroup()


# Join our ukaryotic and prokaryotic blast search results to calculate the HGT index
file3 <-
  full_join(euk, prok, by = "qseqid") %>% # join our two tables by query sequence ID
  replace_na(list( 
    bitscore_d = 0,
    bitscore_r = 0
  )) %>% # here we replace all null values in recipient and donor bitscores with zeros. This is necessary to assign our 'strong' domain classes.
  mutate(h = bitscore_d - bitscore_r) %>% # calculate HGT index (h)colmean
  mutate(
    domain_assignment = # create domain assignment column
      case_when(
        # strong prokaryote
        bitscore_r == 0 & h >= 30 ~ "prokaryote_strong",
        # strong eukaryote
        bitscore_d == 0 & h <= 0 ~ "eukaryote_strong",
        # prokaryote
        h >= 30 ~ "prokaryote",
        # eukaryote
        h <= 0 ~ "eukaryote",
        # indeterminate
        between(h, 0, 30) ~ "indeterminate"
      )
  )

# append our uniprot and ORF header info to the first dataframe
file3_uniprot <-
  file3 %>%
  separate(
    qseqid,
    sep = ":",
    into = c("chrm", "length_orf"),
    remove = F
  ) %>% # split ORF names
  separate(
    length_orf,
    sep = "_",
    into = c("length_orf", "orf_id"),
    remove = T
  ) %>% # split ORF names 2
  separate(
    sseqid_r,
    sep = "\\|",
    into = c("uniprot_prefix_r", "uniprot_id_r", "product_r"),
    remove = T
  ) %>% # sseqid_r split
  separate(
    sseqid_d,
    sep = "\\|",
    into = c("uniprot_prefix_d", "uniprot_id_d", "product_d"),
    remove = F, 
    fill = "right",
    convert = TRUE
  ) %>% # uniprot hits
  separate(
    sseqid_d,
    sep = "\\.",
    into = c("chrm_bac", "bac_gen_peg"),
    remove = T,
    fill = "right",
    convert = FALSE
  ) %>% # bacterial genome hits
  mutate(
    length_orf = str_replace_all(string = length_orf, pattern = "[/(/)$]", replacement = "")
  ) %>% # remove () from orf length
  mutate(chrm = str_replace_all(
    string = chrm,
    pattern = "_quiver_pilon$",
    replacement = ""
  )) %>% # remove quiver pilon from chrm
  left_join(
    .,
    uniprot_reviewed,
    by = c("uniprot_id_r" = "entry")
  ) %>% # Join Uniprot info to recipient hits
  select(-c(starts_with(
    "uniprot_prefix"
  ))) %>%
  dplyr::rename(
    entry_name_uniprot_r = entry_name,
    gene_ontology_biological_process_uniprot_r = gene_ontology_biological_process,
    gene_ontology_molecular_function_uniprot_r = gene_ontology_molecular_function,
    gene_names_uniprot_r = gene_names,
    organism_uniprot_r = organism,
    taxonomic_lineage_superkingdom_r = taxonomic_lineage_superkingdom
  ) %>%
  left_join(
    .,
    uniprot_reviewed,
    by = c(
      "uniprot_id_d" = "entry"
    )
  ) %>% # Join Uniprot info to donor hits
  dplyr::rename(
    entry_name_uniprot_d = entry_name,
    gene_ontology_biological_process_uniprot_d = gene_ontology_biological_process,
    gene_ontology_molecular_function_uniprot_d = gene_ontology_molecular_function,
    gene_names_uniprot_d = gene_names,
    organism_uniprot_d = organism,
    taxonomic_lineage_superkingdom_d = taxonomic_lineage_superkingdom
  ) %>%
  left_join(
    .,
    ORF_headers,
    by = c("qseqid", "chrm", "length_orf", "orf_id")
  ) %>% # join ORF information
  left_join(
    .,
    bac_gff,
    by = c("chrm_bac","bac_gen_peg")
  ) %>% # join our gff info
  as.data.frame() %>% 
  mutate(
    taxonomic_lineage_superkingdom_d = ifelse(
      is.na(
        taxonomic_lineage_superkingdom_d
      ),
      "Bacteria",
      taxonomic_lineage_superkingdom_d
    )
  ) %>% # replace NAs in superkingdom column with bacteria for our bacterial genome hits
  mutate(
    organism_uniprot_d = # assign taxonomy of our bacterial genomes tp organism_uniprot_d column
      case_when(
        # strong prokaryote
        chrm_bac == "0F" ~ "Phycisphaerae sp.",
        chrm_bac == "1F" ~ "Rhodovibrionaceae sp.",
        chrm_bac == "2F" ~ "Ekhidna sp.",
        chrm_bac == "3F" ~ "Balneola sp.",
      )
  ) 
# get a list of uniprot IDs in chrm_bac for removal
remove_these <-
  (file3_uniprot %>%
     select(chrm_bac) %>%
     filter(
       !chrm_bac %in% c(
         "0F",
         "1F",
         "2F",
         "3F"
       )
     ) %>%
     filter(
       !is.na(
         chrm_bac
       )
     )
  )[["chrm_bac"]]

# render non 0-3F entries as NAs
file3_uniprot$chrm_bac[file3_uniprot$chrm_bac %in% remove_these] <- NA

# here we will get the uniprot IDs to download the respective sequences with

# Get habitat for each hit
# export the uniprot IDs first
# file3_uniprot_prok_hits <-
#  file3_uniprot %>%
#   filter(domain_assignment == "prokaryote" | domain_assignment == "prokaryote_strong") %>%
#    select(organism_uniprot_d) %>%
#   filter(!is.na(organism_uniprot_d))

# write_tsv(file3_uniprot_prok_hits, file = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/file3_uniprot_prok_hits.tsv")

# import the updated tsv
file3_uniprot_prok_hits_eco <-
  read_tsv(
    file = here("data","01_psam_intergenic_HGT_index","file3_uniprot_prok_hits_ecology_domain.tsv"),
    col_names = c(
      "organism_uniprot_d",
      "ecology_broad",
      "ecology_fine",
      "domain"
    ),
    skip = 1
  ) %>%
  unique() %>%
  left_join(
    (file3_uniprot %>%
       filter(
         domain_assignment == "prokaryote" | domain_assignment == "prokaryote_strong"
       )),
    .,
    by = "organism_uniprot_d"
  )

# generate just the bac genome hits for our alignment investigation later
file3_uniprot_prok_hits_0Fthru3F <-
  file3_uniprot_prok_hits_eco %>% 
  filter(!is.na(chrm_bac)) %>% 
  mutate(header = paste(
    chrm_bac,
    length_bac,
    sep = ":"
  ))

# here we will have an un-sliced version of the above so that we can isolate all of the best prokarayote hits to our prokaryote/prokaryote_strong h index classifications.
# once we have this, we will download these nucleotide sequences and align them with our qseqid.
prok_unSliced <-
  read_tsv(
    file = here("data","01_psam_intergenic_HGT_index","DMNDblastp_uniprot_output_psam","dmnd_blastp_hgt_output","intergenic_dmnd_uniprot_prok.out.tsv"),
    col_names = colnames_donor,
    progress = T
  ) %>%
  group_by(qseqid) %>% # group our data by sqeqid
  dplyr::filter(
    qseqid %in% file3_uniprot_prok_hits_eco$qseqid
  ) %>%
  ungroup() %>%
  separate(
    sseqid_d,
    sep = "\\|",
    into = c(
      "uniprot_prefix_d",
      "uniprot_id_d",
      "product_d"
    ),
    remove = F
  ) %>% # uniprot hits
  separate(
    sseqid_d,
    sep = ":",
    into = c("chrm_bac", "length_bac"),
    remove = T
  ) %>% # bacterial genome hits
  select(-c(starts_with("uniprot_prefix"))) %>%
  select(uniprot_id_d) %>%
  filter(!is.na(uniprot_id_d)) %>%
  unique()

# write it!
write_delim(
  prok_unSliced,
  file = here("data","file3_uniprot_prok_IDs_all_hits.delim"),
  delim = " ",
  col_names = FALSE
)

# read in out output from uniprot search
uniprot2ncbi_ids_all_hits <-
  read_tsv(
    file = here("data","uniprot-yourlist_M20220411A084FC58F6BBA219896F365D15F2EB444998D7P.tab"),
    col_names = TRUE,
    col_types = NULL
  ) %>%
  janitor::clean_names() %>%
  left_join(
    (read_tsv(
      file = here("data","01_psam_intergenic_HGT_index","DMNDblastp_uniprot_output_psam","dmnd_blastp_hgt_output","intergenic_dmnd_uniprot_prok.out.tsv"),
      col_names = colnames_donor,
      progress = T
    ) %>%
       group_by(qseqid) %>% # group our data by sqeqid
       dplyr::filter(
         qseqid %in% file3_uniprot_prok_hits_eco$qseqid
       ) %>%
       ungroup() %>%
       separate(
         sseqid_d,
         sep = "\\|",
         into = c("uniprot_prefix_d", "uniprot_id_d", "product_d"),
         remove = T
       ) %>% # uniprot hits
       select(-c(starts_with("uniprot_prefix")))),
    .,
    by = c("uniprot_id_d" = "entry")
  ) %>%
  mutate(
    cross_reference_ref_seq = str_extract(
      cross_reference_ref_seq,
      "WP_[:digit:]+.1"
    )
  ) %>%
  rename(
    sseqid = cross_reference_ref_seq
  ) %>%
  dplyr::select(
    -c(
      entry_name,
      uniprot_id_d,
      product_d
    )
  ) %>%
  drop_na() %>%
  relocate(sseqid, .after = qseqid)

write_tsv(
  uniprot2ncbi_ids_all_hits,
  file = here("data","uniprot2ncbi_ids_all_hits.tsv"),
  col_names = FALSE
)

rm(
  euk,
  prok,
  ORF_headers,
  uniprot_reviewed,
  file3,
  colnames_donor,
  colnames_recipients,
  remove_these,
  bac_gff_directory,
  bac_gff_files
)
```

```{r}


#  # X is a sample of 100 normally distributed random variables
# P = ecdf(file3$h)    
# # P is a function giving the empirical CDF of X
# P(0.0)        
#  # This returns the empirical CDF at zero (should be close to 0.5)
# plot(P)

file3_uniprot %>%
  select(
    evalue_r,
    evalue_d,
    bitscore_r,
    bitscore_d,
    h
  ) %>%
  summary()

file3_uniprot %>%
  group_by(domain_assignment) %>%
  summarise(cnt = dplyr::n()) %>%
  mutate(freq = round(cnt / sum(cnt), 3)) %>%
  dplyr::arrange(desc(freq))



# strong prokaryotic blast hits
file3_uniprot_prok_hits_eco %>% 
  filter(bitscore_d >= 100 & evalue_d <= 1e-12) %>% 
  nrow()

# strong prokaryotic blast hits to 0F-3F
file3_uniprot_prok_hits_eco %>% 
  filter(bitscore_d >= 100 & evalue_d <= 1e-12) %>% 
  filter(!is.na(chrm_bac)) %>% 
  nrow()

# strong prokaryotic blast hits to uniprot
file3_uniprot_prok_hits_eco %>% 
  filter(bitscore_d >= 100 & evalue_d <= 1e-12) %>% 
  filter(is.na(chrm_bac)) %>% 
  nrow()

# prokaryotic organism hits for bacteria
file3_uniprot_prok_hits_eco %>% 
  filter(taxonomic_lineage_superkingdom_d == "Bacteria") %>% 
  nrow()

# prokaryotic organism hits for archaea
file3_uniprot_prok_hits_eco %>% 
  filter(taxonomic_lineage_superkingdom_d == "Archaea") %>% 
  nrow()

# 
file3_uniprot_prok_hits_0Fthru3F %>% 
  select(chrm, chrm_bac) %>% 
  table(useNA = "ifany")

# Psammoneis chrm with multiple hits to bacterial genomes
file3_uniprot_prok_hits_0Fthru3F %>% 
  select(chrm, chrm_bac) %>% 
  table(useNA = "ifany") %>% 
  as.data.frame() %>% 
  filter(Freq > 1)

# Determine which of these psammoneis chrmosomes with multiple hits are hitting to the same bacterial genes
file3_uniprot_prok_hits_0Fthru3F %>% 
  filter(chrm %in% c(
    "000002F",
    "000037F",
    "000089F",
    "000103F",
    "000532F"
  )) %>% 
  arrange(chrm) %>% 
  select(
    chrm,
    chrm_bac,
    bac_gen_peg
  ) 

# Now, let's check out the ones that have multiple hits to the same bacterial genes
file3_uniprot_prok_hits_0Fthru3F %>% 
  filter(chrm %in% c(
    "000002F",
    "000089F",
    "000103F",
    "000532F"
  )) %>% 
  arrange(chrm) %>% 
  select(
    chrm,
    chrm_bac,
    bac_gen_peg,
    qstart_d,
    qend_d,
    sstart_d,
    send_d,
    type,
    frame,
    strand,
    bac_gen_gene_annotation
  )


# Seems like 89F is the only one with sufficient evidence for being the result of pseudogenization
file3_uniprot_prok_hits_0Fthru3F %>% 
  filter(chrm == "000089F") %>% 
  select(-ends_with("_r"))

# 
file3_uniprot_prok_hits_0Fthru3F %>% 
  filter(bitscore_d >= 100 & evalue_d <= 1e-12) %>% 
  filter(chrm_bac == "0F") %>%
  filter(chrm == "000002F") %>%
  select(chrm, chrm_bac) %>% 
  table(useNA = "ifany")

# 
file3_uniprot_prok_hits_0Fthru3F %>% 
  filter(bitscore_d >= 100 & evalue_d <= 1e-12) %>% 
  filter(chrm_bac == "0F") %>%
  filter(chrm == "000002F") %>%
  select(chrm, chrm_bac) %>% 
  table(useNA = "ifany")

# ORF length ranges by domain assignment
file3_uniprot %>% 
  select(length, domain_assignment) %>% 
  mutate(length = as.numeric(length)) %>% 
  group_by(domain_assignment) %>% 
  summarise(
    minlength = min(length),  
    maxlength = max(length)
  )
```

We need to visualize and evaluate our HGT index output to determine what *h* value cutoff we should use to determine the HGT index categories.\
Based on the visualization below, an *h* value cutoff of 30 should be fine.

```{r}
#| label: Assess h-value to use as cutoff
#| fig-cap: ECD plot of *h* values
file3_uniprot %>%
  ggplot(aes(x = h)) +
  stat_ecdf(geom = "point") +
  scale_x_continuous(
    breaks = c(seq(-200, 300, by = 50), 30),
    limits = c(-200,300)
  ) +
  xlab("HGT index (<i>h</i>)") +
  ylab("Empirical cumulative distribution of <i>h</i>") +
  theme(
    axis.text.x = element_text(
      angle = 90,
      hjust = 1
    ),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    # panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "none",
    legend.title = element_blank()
  ) +
  geom_vline(
    xintercept = 30,
    color = "black",
    linetype = 3
  )
```

```{r}
#| label: Get uniprot hit info AND find prokaryote overlaps
#| eval: false

# get just our prok contigs and sequences
file3_uniprot_prok <-
  file3_uniprot %>%
  filter(domain_assignment == "prokaryote" | domain_assignment == "prokaryote_strong") %>%
  group_by(chrm) %>% # group by ORF chrm
  arrange(chrm, length_orf) %>% # arrange by chrm, then length_orf w/in each chrm
  separate(col = coords, into = c("startcoord_orf", "stopCoord_orf"), remove = T) %>% # split length_orf into startcoord_orf stopCoord_orf
  # collapse our ranges in each chromosome group
  select(
    startcoord_orf,
    stopCoord_orf,
    strand,
    chrm
  ) %>%
  dplyr::rename("start" = "startcoord_orf", "end" = "stopCoord_orf") %>%
  mutate(start = as.numeric(start)) %>%
  mutate(end = as.numeric(end))


# Use plyranges and genomicranges::reduce to merge all overlapping seqs per chromosome and strand
file3_uniprot_prok_overlaps <-
  file3_uniprot %>%
  filter(domain_assignment == "prokaryote" | domain_assignment == "prokaryote_strong") %>%
  group_by(chrm) %>% # group by ORF chrm
  arrange(chrm, length_orf) %>% # arrange by chrm, then length_orf w/in each chrm
  separate(col = coords, into = c("startcoord_orf", "stopCoord_orf"), remove = T) %>% # split length_orf into startcoord_orf stopCoord_orf
  # collapse our ranges in each chromosome group
  select(startcoord_orf, stopCoord_orf, strand, chrm) %>%
  dplyr::rename("start" = "startcoord_orf", "end" = "stopCoord_orf") %>%
  mutate(start = as.numeric(start)) %>%
  mutate(end = as.numeric(end)) %>%
  plyranges::as_granges(seqnames = chrm) %>%
  GenomicRanges::reduce() %>%
  as.data.frame()

# export these as a bed file
file3_uniprot_prok_overlaps %>%
  dplyr::select(seqnames, start, end) %>%
  write_tsv(., file = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/psam_overlaps.bed", col_names = F, )

# extend seqs by 1k length and then reduce
file3_uniprot_prok_overlaps_1k <-
  file3_uniprot_prok_overlaps %>%
  mutate(start_1k = as.numeric(start) - 1000) %>%
  mutate(stop_1k = as.numeric(end) + 1000)

file3_uniprot_prok_overlaps_1k[file3_uniprot_prok_overlaps_1k < 0] <- 1

file3_uniprot_prok_overlaps_1k <-
  file3_uniprot_prok_overlaps_1k %>%
  select(start_1k, stop_1k, strand, seqnames) %>%
  dplyr::rename("start" = "start_1k", "end" = "stop_1k") %>%
  mutate(start = as.numeric(start)) %>%
  mutate(end = as.numeric(end)) %>%
  plyranges::as_granges(seqnames = seqnames) %>%
  GenomicRanges::reduce() %>%
  as.data.frame()


# maybe get aa bed file of overlaps that are in the duplicate hits list below

# export these as a bed file
write_tsv()

# see how many overlaps there are
inner_join(as.data.frame(table(file3_uniprot_prok$chrm)),
           as.data.frame(table(file3_uniprot_prok_overlaps$seqnames)),
           by = "Var1"
) %>%
  inner_join(., as.data.frame(table(file3_uniprot_prok_overlaps_1k$seqnames)), by = "Var1") %>%
  mutate(n_overlaps = Freq.x - Freq.y) %>%
  mutate(n_overlaps_1k = Freq.x - Freq) %>%
  summarise(
    total = sum(Freq.x),
    sum_nonOverlap = sum(Freq.x) - sum(n_overlaps),
    sum_nonOverlap_1k = sum(Freq.x) - sum(n_overlaps_1k),
    sum_overlaps = sum(n_overlaps),
    sum_overlaps_1k = sum(n_overlaps_1k)
  ) %>%
  pivot_longer(cols = everything()) %>%
  ggplot(aes(x = name, y = value, fill = name)) + # Plot with values on top
  geom_bar(stat = "identity") +
  geom_text(aes(label = value), vjust = -0.2) +
  viridis::scale_color_viridis(discrete = T) +
  viridis::scale_fill_viridis(discrete = T) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.5),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "right",
    legend.title = element_blank()
  )
```

```{r}
#| label: prokaryotic hits to the same gene/organism

# here we will get the uniprot IDs to download the respective sequences with
file3_uniprot_prok_IDs <-
  file3_uniprot %>%
  filter(domain_assignment == "prokaryote" | domain_assignment == "prokaryote_strong") %>%
  select(entry_name_uniprot_d) %>%
  filter(!is.na(entry_name_uniprot_d)) %>%
  distinct()

# write the uniprto IDs into a space-delimited file
write_delim(file3_uniprot_prok_IDs, file = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/file3_uniprot_prok_IDs.delim", delim = " ", col_names = FALSE)

# after we search for the above IDs on uniprot, we need to read this in to filter out just WP ncbi IDs to search for, not all have them so the DF is reduced
uniprot2ncbi_ids <-
  read_tsv(file = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/uniprot-yourlist_M202204014ABAA9BC7178C81CEBC9459510EDDEA34887D7H.tab", col_names = TRUE, col_types = NULL) %>%
  janitor::clean_names() %>%
  mutate(cross_reference_ref_seq = str_extract(cross_reference_ref_seq, "WP_[:digit:]+.1")) %>%
  drop_na() %>%
  select(cross_reference_ref_seq)

write_delim(uniprot2ncbi_ids, file = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/uniprot2ncbi_ids.delim", delim = " ", col_names = FALSE)

# now you can just export those and search them via https://www.uniprot.org/uploadlists/ to get the sequences info/fasta

# Determine if any of the overlapping sequences from each chromosome hit to the same thing

# here we will get all chromosomes that have hits to the same uniprot ID
file3_uniprot_prok_IDs_dupes <-
  file3_uniprot %>%
  filter(domain_assignment == "prokaryote" | domain_assignment == "prokaryote_strong") %>%
  select(entry_name_uniprot_d, chrm) %>%
  filter(!is.na(entry_name_uniprot_d)) %>%
  janitor::get_dupes()

# here we will get all dupes to our bacteria
file3_chrm_bac_dupes <-
  file3_uniprot %>%
  filter(domain_assignment == "prokaryote" | domain_assignment == "prokaryote_strong") %>%
  select(chrm_bac, chrm) %>%
  filter(!is.na(chrm_bac)) %>%
  janitor::get_dupes()
```

```{r}
#| label: prokaryote overlap visualizations
#| eval: false

# visualize
ir <-
  IRanges(
    start = file3_uniprot_prok_overlaps$start,
    end = file3_uniprot_prok_overlaps$end
  )
bins <-
  disjointBins(IRanges(start(ir), end(ir) + 1))

dat <-
  cbind(as.data.frame(file3_uniprot_prok_overlaps), bin = bins)

ir_1k <-
  IRanges(
    start = file3_uniprot_prok_overlaps_1k$start,
    end = file3_uniprot_prok_overlaps_1k$end
  )

bins_1k <-
  disjointBins(IRanges(start(ir_1k), end(ir_1k) + 1))

dat_1k <-
  cbind(as.data.frame(file3_uniprot_prok_overlaps_1k), bin = bins_1k)

# visualize density of lengths
file3_uniprot_prok_overlaps %>%
  # filter(seqnames == "000015F") %>%
  ggplot() +
  geom_density_ridges(aes(x = width, y = strand))

# visualize density of lengths
file3_uniprot_prok_overlaps %>%
  # filter(seqnames == "000015F") %>%
  ggplot(aes(x = width, y = strand)) +
  stat_slab(aes(thickness = stat(pdf*n)), scale = 0.7) +
  stat_dotsinterval(side = "bottom", scale = 0.7, slab_size = NA) 

p_dat <-
  dat %>%
  # filter(seqnames == "000015F") %>%
  ggplot() +
  geom_rect(aes(
    xmin = start, xmax = end,
    ymin = bin, ymax = bin + 0.9, fill = seqnames, color = seqnames
  )) +
  viridis::scale_color_viridis(discrete = T) +
  viridis::scale_fill_viridis(discrete = T) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "none",
    legend.title = element_blank()
  )

p_dat_1k <-
  dat_1k %>%
  # filter(seqnames == "000015F") %>%
  ggplot() +
  geom_rect(aes(
    xmin = start, 
    xmax = end,
    ymin = bin, 
    ymax = bin + 0.9, 
    fill = seqnames, 
    color = seqnames
  )) +
  viridis::scale_color_viridis(discrete = T) +
  viridis::scale_fill_viridis(discrete = T) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "none",
    legend.title = element_blank()
  )

p_dat + p_dat_1k
```

```{r}
#| label: prokaryote hit visualizations
#| eval: false

file3_uniprot %>%
  select(domain_assignment, starts_with("length_")) %>%
  pivot_longer(cols = c("length_r", "length_d")) %>%
  mutate(name = factor(name, levels = c("length_r", "length_d"))) %>%
  ggplot(aes(y = domain_assignment, x = value, color = name)) +
  geom_density_ridges(alpha = 0.5) +
  viridis::scale_color_viridis(discrete = T) +
  viridis::scale_fill_viridis(discrete = T) +
  scale_x_continuous(breaks = c(seq(0, 1500, by = 100))) +
  xlab("Blastp hit length") +
  ylab("") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "right",
    legend.title = element_blank()
  )


file3_uniprot %>%
  select(domain_assignment, starts_with("length_")) %>%
  pivot_longer(cols = c("length_r", "length_d")) %>%
  mutate(name = factor(name, levels = c("length_r", "length_d"))) %>%
  ggplot(aes(x = value, color = name)) +
  viridis::scale_color_viridis(discrete = T) +
  viridis::scale_fill_viridis(discrete = T) +
  stat_ecdf(geom = "point", size = 0.75, alpha = 0.5) +
  facet_wrap(vars(domain_assignment), scales = "free") +
  # scale_x_continuous(breaks = c(seq(0, 1500, by = 100))) +
  xlab("Blastp hit length") +
  ylab("") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "right",
    legend.title = element_blank()
  )



file3_uniprot_prok_hits_eco %>%
  tidyr::replace_na(list(domain = "Bacteria")) %>%
  select(domain) %>%
  table(useNA = "ifany") %>%
  as.data.frame() %>%
  ggplot(aes(x = ., y = Freq, fill = .)) + # Plot with values on top
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = -0.15) +
  viridis::scale_color_viridis(discrete = T) +
  viridis::scale_fill_viridis(discrete = T) +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "right",
    legend.title = element_blank()
  )


file3_uniprot_prok_hits_eco %>%
  tidyr::replace_na(list(ecology_broad = "Marine")) %>%
  select(ecology_broad) %>%
  table(useNA = "ifany") %>%
  as.data.frame() %>%
  ggplot(aes(x = ., y = Freq, fill = .)) + # Plot with values on top
  geom_bar(stat = "identity") +
  geom_text(aes(label = Freq), vjust = -0.15) +
  viridis::scale_color_viridis(discrete = T) +
  viridis::scale_fill_viridis(discrete = T) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 0.5),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "none",
    legend.title = element_blank()
  )
```

```{r}
#| label: sequence similarity ordination prep
#| eval: false

# first, read in ORFs and filter to only those with prokaryote hits
ORFs <-
  read.fasta(file = here("data","01_psam_intergenic_HGT_index","psam_genome","orfipy_psamm_complement_intergenic.fa_out","psamm_complement_ORFs_orfipy.fa"), as.string = T, seqtype = "DNA")

psamm_cds <-
  read.fasta(here("data","01_psam_intergenic_HGT_index","DMNDblastp_uniprot_output_psam", "01_genomes_CDS"), as.string = T, seqtype = "DNA")

# subsample psam CDS genes
psamm_subsampled <-
  as.data.frame(do.call(rbind, psamm_cds)) %>%
  slice_sample(prop = 0.015) %>%
  rownames_to_column(var = "V2") %>%
  mutate(V2 = paste(">", V2, sep = "")) %>% # add the carrot to make it the fasta header
  mutate(V1 = str_replace(V1, "[:punct:]$", "")) # remove the * cstop character so that muscle dowsn't throw a fit

write.table(psamm_subsampled,
  sep = "\n",
  file = here("data","01_psam_intergenic_HGT_index","sequence_ordination","psam_genic_subsampled.fa"),
  quote = FALSE, row.names = FALSE, col.names = FALSE
)

# filter those seqs
ORFs_filt <-
  as.data.frame(do.call(rbind, ORFs)) %>%
  rownames_to_column(var = "V2") %>%
  dplyr::filter(V2 %in% (file3_uniprot %>%
    filter(domain_assignment == "prokaryote" | domain_assignment == "prokaryote_strong") %>%
    pull(qseqid))) %>%
  mutate(V2 = paste(">", V2, sep = "")) %>% # add the carrot to make it the fasta header
  mutate(V1 = str_replace(V1, "[:punct:]$", "")) # remove the * cstop character so that muscle dowsn't throw a fit

write.table(ORFs_filt[, 1:2],
  sep = "\n",
  file = here("data","01_psam_intergenic_HGT_index","sequence_ordination","psamm_prokaryote_candidate_ORFs_filtered.fa"),
  quote = FALSE, row.names = FALSE, col.names = FALSE
)

# read in prtein version of ORFs
ORFs_AA <-
  read.fasta(file = here("data","01_psam_intergenic_HGT_index","psam_genome","orfipy_psamm_complement_intergenic.fa_out","psamm_complement_ORFs_orfipy.faa"), as.string = T, seqtype = "AA")

# filter those seqs to get jsut our prokaryote HGT candidates
ORFs_AA_filt <-
  as.data.frame(do.call(rbind, ORFs_AA)) %>%
  rownames_to_column(var = "V2") %>%
  dplyr::filter(V2 %in% (file3_uniprot %>%
    filter(domain_assignment == "prokaryote" | domain_assignment == "prokaryote_strong") %>%
    pull(qseqid))) %>%
  mutate(V2 = paste(">", V2, sep = "")) %>% # add the carrot to make it the fasta header
  mutate(V1 = str_replace(V1, "[:punct:]$", "")) # remove the * cstop character so that muscle dowsn't throw a fit

# write fasta of proteins for final ncbi blastp search
write.table(ORFs_AA_filt[, 1:2],
  sep = "\n",
  file = here("data","01_psam_intergenic_HGT_index","ncbi_blastp_final","psamm_prokaryote_hgt_candidate_ORFs.faa"),
  quote = FALSE, row.names = FALSE, col.names = FALSE
)


# Read in the NCBI BlastP results
ORF_ncbi_blastp <-
  read_tsv(
    here("data","01_psam_intergenic_HGT_index","ncbi_blastp_final","ncbi_blastp_output","Psammoneis_japonica_classA_seqs_blastp_out.tsv"),
    col_names = c("qseqid", "sseqid", "pident", "length", "mismatch", "gapopen", "qstart", "qend", "sstart", "send", "evalue", "bitscore", "sacc", "staxids", "sscinames"), 
    progress = T) %>% 
  mutate(acc.ver = str_remove_all(sseqid, paste(c("^[:alpha:]+\\|+","\\|$","\\|[:alpha:]$"), collapse="|"))) 


# open our arrow datasets
prot.accession2taxid <- 
  open_dataset("/Users/cbg/local/ncbi/prot.accession2taxid.FULL_2022-10-09_parquet")


# read in taxid categories file
taxid_categories <-
  read_tsv(
    file = "/Users/cbg/local/ncbi/categories.dmp",
    col_names = c("category", "taxid_species_level", "taxid"), 
    progress = T) %>% 
  mutate(category_expanded = case_when(
  category == "A" ~ "Archaea",
  category == "B" ~ "Bacteria",
  category == "E" ~ "Eukaryota",
  category == "V" ~ "Viruses and Viroids",
  category == "U" ~ "Unclassified",
  category == "O" ~ "Other"
  ))


# join our datasets
ORF_blastp_results <- 
  left_join( # join our blastp results with our prot.accession2taxid data from the parquet files
    ORF_ncbi_blastp,
    (prot.accession2taxid %>% 
  filter(accession.ver %in% ORF_ncbi_blastp$acc.ver) %>% 
  collect()),
  by = c("acc.ver" = "accession.ver")
  ) %>% 
  left_join(., taxid_categories, by = "taxid") # Join our taxid_categories file to the above

  
ORF_blastp_results_joined <- 
  left_join(ORF_blastp_results,
            (ORF_blastp_results %>% 
               group_by(qseqid) %>% 
               select(qseqid, category_expanded) %>% 
               table() %>% 
               as.data.frame() %>% 
               pivot_wider(names_from = category_expanded, values_from = Freq) %>% 
               mutate( # now we make a column to calculate the total hits to each qseqid
                 total_hits = rowSums(select(., contains(c("Arc","Bac","Euk"))))
                 )),
            by = "qseqid") %>% 
  group_by(qseqid) %>% 
  slice_max(bitscore, n = 1, with_ties = FALSE) %>% 
  ungroup() %>% 
  left_join(., ORF_headers, by = "qseqid") %>% 
  filter(category == "B")

ORF_blastp_results_joined %>% 
  filter(!qseqid %in% file3_uniprot_prok_hits_0Fthru3F$qseqid) %>% 
  dplyr::count(category)

test <-
  ORF_blastp_results_joined %>% 
  filter(qseqid %in% file3_uniprot_prok_hits_0Fthru3F$qseqid)

# rm() objects that aren't needed
rm(ORFs, ORFs_AA, ORFs_filt, ORFs_AA_filt, psamm_cds, psamm_subsampled)

# read in the filtered and subsampled files as codon tables
ORFs <-
  codonTable(readSet(file = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/sequence_ordination/psam_genic_subsampled.fa", ))

psamm_cds <-
  codonTable(readSet(file = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/sequence_ordination/psamm_prokaryote_candidate_ORFs_filtered.fa"))

uniprot_nt <-
  codonTable(readSet(file = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/sequence_ordination/uniprot_NT_seqs/uniprot_NT_seqs.fa"))
```

We can see from the graphs and stats below, that the nucleotide sequences of the ORFs and the Psamm CDS sequences display a greater similarity to each other than the ORFs do to the nucleotide sequences of the UniProt nucleotide sequences. Based on this, it seems that the potential HGT candidates were transferred long ago and have started to take on the characteristics of the recipient genome. We will need to download nucleotide sequences for all blast hits and align them to the ORFs to determine what mutations have occurred.

```{r}
#| label: Codon usage comparison between uniprot hits
#| eval: false

intraBplot(ORFs, psamm_cds,
  names = c("ORFs", "psamm_cds"), variable = "MILC", ribosomal = FALSE,
  size = 1, alpha = 0.5
)

intraBplot(ORFs, uniprot_nt,
  names = c("ORFs", "uniprot"), variable = "MILC", ribosomal = FALSE,
  size = 1, alpha = 0.5
)

intraBplot(psamm_cds, uniprot_nt,
  names = c("psamm_cds", "uniprot"), variable = "MILC", ribosomal = FALSE,
  size = 1, alpha = 0.5
)


ORFs_MILC <-
  as.data.frame(MILC(ORFs)) %>%
  rename(milc = self) %>%
  add_column(group = "ORF")

psamm_cds_MILC <-
  as.data.frame(MILC(psamm_cds)) %>%
  rename(milc = self) %>%
  add_column(group = "Psammoneis_CDS")

uniprot_nt_MILC <-
  as.data.frame(MILC(uniprot_nt)) %>%
  rename(milc = self) %>%
  add_column(group = "uniprot")

dat <-
  rbind(ORFs_MILC, psamm_cds_MILC, uniprot_nt_MILC)

boxplot(milc ~ group, data = dat)


milc_aov <-
  aov(milc ~ factor(group), data = dat)

summary(milc_aov)
report(milc_aov)

TukeyHSD(milc_aov, conf.level = .95)
broom::tidy(TukeyHSD(milc_aov, conf.level = .95))

plot(milc_aov)

# convert codontables to df
codon_counts <-
  rbind(
    (as.data.frame(codonCounts(object = ORFs)) %>%
      add_column(seqs = ORFs@ID) %>%
      add_column(color = "blue")),
    (as.data.frame(codonCounts(object = psamm_cds)) %>%
      add_column(seqs = psamm_cds@ID) %>%
      add_column(color = "red")),
    (as.data.frame(codonCounts(object = uniprot_nt)) %>%
      add_column(seqs = uniprot_nt@ID) %>%
      add_column(color = "green"))
  )

colors <-
  codon_counts$color


# CCA
ca <- FactoMineR::CA(X = (codon_counts %>%
  select(-c(color, seqs))))

FactoMineR::plot.CA(ca, label = "row.sup", col.row = colors, col.col = "black", graph.type = "ggplot", new.plot = TRUE)

legend(
  bty = "y", x = "topright",
  legend = c("CDS", "ORFs", "Bacteria"),
  fill = c("red", "blue", "green"),
  border = "black"
)
```

To obtain the nucleotide sequences for our UniProt hits we will need to search the protein accessions in genbank. Then download the GI and summary files. We grep the Accession + GI \# lines out of the summary to join with our late files. We then use the GI number file as input to [NCBI Batch Entrez](https://www.ncbi.nlm.nih.gov/sites/batchentrez) and search against the Identical Protein Groups (IPG) portion of NCBI. We then download the IPG file for our GI hits. From here we read in our files to R and join them based on GI and protein accessions. After that we can pull just the accessions for all nucleotide sequences for our AA sequences and then use those to align our ORFs against.

```{r}
#| label: UniProt hits IPG files
#| eval: false


# read in our fasta file names (named after the protein sequence the nuclotide sequence translates into)
NT_seq_files <-
  list.files(path = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/hgt_candidates_alignments/nt_seqs/", full.names = F) 

NT_seq_paths <-
  list.files(path = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/hgt_candidates_alignments/nt_seqs", full.names = T) 

NT_files_paths <-
  as.data.frame(cbind(NT_seq_files, NT_seq_paths))

for (n in 1:nrow(NT_files_paths)) {
  df <- data.frame()
  {
    df <- rbind(df,
    as.data.frame(do.call(rbind, (read.fasta(file = NT_files_paths$NT_seq_paths[n],
           seqtype = "DNA", as.string = TRUE, whole.header = T))))  %>% 
    rownames_to_column(var = "header")
    )
    }
  NT_files_paths_seqs <-
    cbind(NT_files_paths, df)
}

nt_seqs_df <-
  as.data.frame(NT_files_paths_seqs) %>% 
  mutate(sseqid = str_remove(NT_seq_files, "\\.fasta"))

ORF_clstrs <-
  left_join(uniprot2ncbi_ids_all_hits, nt_seqs_df, by = 'sseqid') %>% 
  rename(sseqid_seq = V1)


# use a for loop to recursively create directories for each ORF alignment
for (n in 1:length(unique(ORF_clstrs$qseqid))) {
  dir <- "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/hgt_candidates_alignments/alignments/"
  {
    dir.create(path = paste(dir, str_replace(str_remove(unique(ORF_clstrs$qseqid)[n], "\\(\\)"), "\\:", "_"), sep = ""))
  }}

# Here we coerce our bacterial genome hist into bed files for each genome
bed_0F <-
  file3_uniprot %>% 
  filter(chrm_bac == "000000F") %>% 
  select(chrm_bac, length_bac) %>% 
  separate(length_bac, into = c("chromStart","chromEnd"), remove = T, extra = "merge") %>% 
  add_column("name" = "0F", "score" = 0) %>%  
  mutate(strand = str_extract(chromEnd, "\\+|\\-")) %>% 
  mutate(chromEnd = str_remove(chromEnd, "\\(\\+\\)|\\(\\-\\)")) %>% 
  unique()
  

bed_1F <-
  file3_uniprot %>% 
  filter(chrm_bac == "000001F") %>% 
  select(chrm_bac, length_bac) %>% 
  separate(length_bac, into = c("chromStart","chromEnd"), remove = T, extra = "merge") %>% 
  add_column("name" = "1F", "score" = 0) %>%  
  mutate(strand = str_extract(chromEnd, "\\+|\\-")) %>% 
  mutate(chromEnd = str_remove(chromEnd, "\\(\\+\\)|\\(\\-\\)")) %>% 
  unique()
  

bed_2F <-
  file3_uniprot %>% 
  filter(chrm_bac == "000002F") %>% 
  select(chrm_bac, length_bac) %>% 
  separate(length_bac, into = c("chromStart","chromEnd"), remove = T, extra = "merge") %>% 
  add_column("name" = "2F", "score" = 0) %>%  
  mutate(strand = str_extract(chromEnd, "\\+|\\-")) %>% 
  mutate(chromEnd = str_remove(chromEnd, "\\(\\+\\)|\\(\\-\\)")) %>% 
  unique()
  

bed_3F <-
  file3_uniprot %>% 
  filter(chrm_bac == "000003F") %>% 
  select(chrm_bac, length_bac) %>% 
  separate(length_bac, into = c("chromStart","chromEnd"), remove = T, extra = "merge") %>% 
  add_column("name" = "3F", "score" = 0) %>%  
  mutate(strand = str_extract(chromEnd, "\\+|\\-")) %>% 
  mutate(chromEnd = str_remove(chromEnd, "\\(\\+\\)|\\(\\-\\)")) %>% 
  unique()
  


# Now we write the bed files and use bedtools getfasta to extract them
write_tsv(bed_0F, 
          "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/hgt_candidates_alignments/0F_prok_hits.bed", 
          col_names = F)
write_tsv(bed_1F, 
          "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/hgt_candidates_alignments/1F_prok_hits.bed", 
          col_names = F)
write_tsv(bed_2F, 
          "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/hgt_candidates_alignments/2F_prok_hits.bed", 
          col_names = F)
write_tsv(bed_3F, 
          "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/hgt_candidates_alignments/3F_prok_hits.bed", 
          col_names = F)

# Next we use bedtools getfasta to extract the sequences from each genome
# We use the following to do that: 
# $ bedtools getfasta -fi <input FASTA> -bed <BED> -name -s

# bedtools getfasta -fi /Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/psam_bac_genomes/000000F.quiver_pilon.n_f.fasta -bed 0F_prok_hits.bed -fo 0F_prok_hits.fa -s
# bedtools getfasta -fi /Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/psam_bac_genomes/000001F.quiver_pilon.n_f.fasta -bed 1F_prok_hits.bed -fo 1F_prok_hits.fa -s
# bedtools getfasta -fi /Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/psam_bac_genomes/000002F.quiver_pilon.n_f.fasta -bed 2F_prok_hits.bed -fo 2F_prok_hits.fa -s
# bedtools getfasta -fi /Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/psam_bac_genomes/000003F.quiver_pilon.n_f.fasta -bed 3F_prok_hits.bed -fo 3F_prok_hits.fa -s

# read in fasta output files, we use seqinr::rad.fasta becasue we can choose to keep the DNA as a string, rather than individual characters. 
# We wrap read.fasta into as.data.frame(do.call(rbind, read.fasta())) so that it reads in the fasta as the seqfastadna list and converts it into a dataframe by binding its rows.
prok_hits_seqs_0F <- 
  as.data.frame(do.call(rbind, (read.fasta(file = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/hgt_candidates_alignments/0F_prok_hits.fa",
           seqtype = "DNA", as.string = TRUE, whole.header = T)))) %>% 
  rownames_to_column(var = "header")

prok_hits_seqs_1F <- 
  as.data.frame(do.call(rbind, (read.fasta(file = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/hgt_candidates_alignments/1F_prok_hits.fa",
           seqtype = "DNA", as.string = TRUE, whole.header = T)))) %>% 
  rownames_to_column(var = "header")

prok_hits_seqs_2F <- 
  as.data.frame(do.call(rbind, (read.fasta(file = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/hgt_candidates_alignments/2F_prok_hits.fa",
           seqtype = "DNA", as.string = TRUE, whole.header = T)))) %>% 
  rownames_to_column(var = "header")

prok_hits_seqs_3F <- 
  as.data.frame(do.call(rbind, (read.fasta(file = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/hgt_candidates_alignments/3F_prok_hits.fa",
           seqtype = "DNA", as.string = TRUE, whole.header = T)))) %>% 
  rownames_to_column(var = "header")

sseqid_0_3F <-
  rbind(prok_hits_seqs_0F, prok_hits_seqs_1F, prok_hits_seqs_2F, prok_hits_seqs_3F) %>% 
  left_join(file3_uniprot_prok_hits_0Fthru3F, ., by = 'header') %>% 
  dplyr::rename(sseqid_seq = V1)

# Join and drop duplicates between ncbi hits and hits to our bacterial genomes
sseqid_sequences <-
  full_join(sseqid_0_3F, ORF_clstrs, by = c('qseqid', 'header', 'sseqid_seq'))

# read in our ORFs for filtering
ORF_seqs <-
  as.data.frame(do.call(rbind, (read.fasta(
    file = "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/psam_genome/orfipy_psamm_complement_intergenic.fa_out/psamm_complement_ORFs_orfipy.fa", 
    seqtype = "DNA", as.string = TRUE, whole.header = T)))) %>% 
  rownames_to_column(var = "header") %>% 
  separate(header, into = c("qseqid", "coords_strand", "type", "length", "frame", "start", "stop"), sep = " ") %>%
  separate(qseqid, sep = ":", into = c("chrm", "length_orf"), remove = F) %>% # split and clean qseqid
  separate(length_orf, sep = "_", into = c("length_orf", "orf_id"), remove = T) %>% # split ORF names 2
  mutate(length_orf = str_replace_all(string = length_orf, pattern = "[/(/)$]", replacement = "")) %>% # remove () from orf length
  mutate(chrm = str_replace_all(string = chrm, pattern = "_quiver_pilon$", replacement = "")) %>% # remove quiver pilon from chrm
  mutate(strand = str_extract_all(coords_strand, "\\(\\+\\)|\\(\\-\\)")) %>%
  mutate(strand = str_replace_all(string = strand, pattern = "^\\(|\\)$", replacement = "")) %>% # split coords_strand
  mutate(coords_strand = str_replace_all(string = coords_strand, pattern = "^\\[|\\]|\\(|\\+\\)$|\\-\\)$", replacement = "")) %>% # split coords_strand
  dplyr::rename(coords = coords_strand) %>% 
  # left_join(uniprot2ncbi_ids_all_hits, ., by = "qseqid") %>% 
  rename(qseqid_seq = V1)

#
qseqid_and_sseqid_sequences <-
  left_join(sseqid_sequences, ORF_seqs, by = 'qseqid') %>% 
  dplyr::select(qseqid, header, qseqid_seq, sseqid_seq) %>% 
  filter(!is.na(sseqid_seq))



# now we need to write the individual fasta files for each ORF and its respective hit sequences to their respective folders.

for (n in 1:length(unique(qseqid_and_sseqid_sequences$qseqid))){
  print(n)
  
  df_qseq <-
    reshape2::melt(qseqid_and_sseqid_sequences, id = c("qseqid", "header")) %>% 
    dplyr::filter(qseqid == unique(qseqid_and_sseqid_sequences$qseqid)[n], variable == "qseqid_seq") %>% 
    dplyr::select(header, value) %>% 
    mutate(header = paste(">", header, sep = ""))

  df_sseq <-
    reshape2::melt(qseqid_and_sseqid_sequences, id = c("qseqid", "header")) %>% 
    dplyr::filter(qseqid == unique(qseqid_and_sseqid_sequences$qseqid)[n], variable == "sseqid_seq") %>% 
    dplyr::select(qseqid, value) %>% 
    dplyr::rename(header = qseqid) %>% 
    mutate(header = paste(">", header, sep = ""))

  df_combined <- rbind(df_qseq,df_sseq)
  
write.table(df_combined,
  sep = "\n",
  file = paste("/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/01_psam_intergenic_HGT_index/hgt_candidates_alignments/alignments/",str_replace(str_remove(unique(qseqid_and_sseqid_sequences$qseqid)[n], "\\(\\)"), "\\:", "_"),"_qseq_sseq.fa", sep = ""),
  quote = FALSE, row.names = FALSE, col.names = FALSE)
}

# Run our alignment loop in the terminal:
#  for i in *.fa; do muscle -align $i -output $i.afa; done

```

# HGT index of genic orthogroups (**h**<sub>orth</sub>)

First, Wade Roberts obtained proteomes for Psammoneis and 18 other taxa:\
Psammoneis japonica\
Pseudo-nitzschia multiseries\
Pseudo-nitzschia multistriata\
Fragilariopsis cylindrus\
Seminavis robusta\
Fistulifera solaris\
Phaeodactylum tricornutum\
Nitzschia sp_Nitz4\
Cyclotella cryptica\
Cyclotella nana\
Nannochloropsis gaditana and Ectocarpus siliculosus as outgroups

Wade Roberts then ran Orthofinder on these proteomes using default settings to produce orthologs.

Then, I sorted files to include only lines for orthologs with 'Psammoneis' genes present. I performed this on a concatenated version of the orthogroups and unassigned gene files so that I would also have singleton genes in there.  

    cat Orthogroups.tsv <(tail -n+2 Orthogroups_UnassignedGenes.tsv) | grep 'Psammoneis' | > Orthogroups_Psam_filtered.tsv

Then I copy the .fa files for those orthogroups to their own folder.

    mkdir Orthogroup_Sequences_Psammoneis
    cd Orthogroup_Sequences
    cp $(ls | grep "$(awk '{ print $1 }' ../Orthogroups/Orthogroups_Psam_filtered.tsv)" - ) ../Orthogroup_Sequences_Psammoneis/

To perform the HGT index blasts, I used Diamond blastp in a for loop over each orthogroup .fa file. This portion of the analysis was performed on the AHPCC computing cluster using the slurm file 'submit_diamond_HGT_index_768gb.slurm '.\
Our reference databases are the same as those used in the intergenic HGT index above.

We extract and translate the CDS sequences for our bacterial genomes in the same method as in our intergenic HGT analysis.

    Rscript gff_filter_rename_bac.R

    bedtools getfasta -fi 000000F.quiver_pilon.n_f.fasta -bed 0F.gff -fo 000000F.fa -s
    bedtools getfasta -fi 000001F.quiver_pilon.n_f.fasta -bed 1F.gff -fo 000001F.fa -s
    bedtools getfasta -fi 000002F.quiver_pilon.n_f.fasta -bed 2F.gff -fo 000002F.fa -s
    bedtools getfasta -fi 000003F.quiver_pilon.n_f.fasta -bed 3F.gff -fo 000003F.fa -s

    echo 'translate bacterial genome sequences'
    Rscript translate_cds.R

We construct our diamond blast databases in the same manner as in our intergenic HGT analysis.

    ## Here we construct our Diamond blast database
    # append our CDS AA sequences to our .fna file for blast input
    echo 'append our CDS AA sequences to our .faa file for blast input'
    cat 00000*F_gff.faa >> uniprot_reviewed_Arc_2157_Bac_2.fasta


    echo 'rm initial bacterial genome files'
    rm 00000*F.fa
    rm psam.final_assembly.Quiver_Pilon_Reapr_GapCloser.fa

    # make diamond db
    echo 'Make Diamond DB using NR + 0-3F bacterial genomes'
    diamond makedb --in uniprot_reviewed_Arc_2157_Bac_2.fasta -d uniprot_prok
    diamond makedb --in uniprot-reviewed_yes+taxonomy_2759_eukaryota.fasta -d uniprot_euk

    mkdir HGT_index_output_donor/
    mkdir HGT_index_output_recipient/
    mkdir HGT_index_output/

Now we use the bash loop file to perform our blast searches. This script runs two loops, one loop searches each orthogroup against the donor (Archaea & Bacteria) diamond blast database and the other loop searches against the recipient (Eukaryote) blast database. At the end of each loop, the respective output .tsv files are moved into their respective output directories.

    # Start our Diamond blast bash loop
    echo 'Start the diamond blast for loop.'
    ./diamond_blast_orthogroups_768gb.sh

Finally, we enter the directory for both donor and recipient output files and append a file name column to each output file. This column contains the name of the input file so that we can keep track of the output from each orthogroup. We then add column names as the first line of every output file.

    cd HGT_index_output_donor/
    for f in *.tsv; do sed -i "s/$/\t$f/" $f; done
    for f in *.tsv; do sed -i '1 s/^/qseqid\tsseqid_d\tpident_d\tlength_d\tmismatch_d\tgapopen_d\tqstart_d\tqend_d\tsstart_d\tsend_d\tevalue_d\tbitscore_d\torthogroup\n/' $f; done

    cd ../HGT_index_output_recipient/
    for f in *.tsv; do sed -i "s/$/\t$f/" $f; done
    for f in *.tsv; do sed -i '1 s/^/qseqid\tsseqid_r\tpident_r\tlength_r\tmismatch_r\tgapopen_r\tqstart_r\tqend_r\tsstart_r\tsend_r\tevalue_r\tbitscore_r\torthogroup\n/' $f; done

## *h*<sub>orth</sub> calculation


```{r}
#| label: Crisp HGT index classification

# HGT index

# calculating hgt index (h)
# From Boschettis et al. 2012:
#   The HGT index (h) is calculated as the difference between the highest non-metazoan [donor] and the highest metazoan [recipient] bitscore.

# To calculate HGT classes A, B, & C. (C,(B,(A))), as per Crisp et al. 2015.
# Class C: h >= 30 & best donor bitscore >= 100
# Class B: For each class C gene, the average h value of the members of its ortholog group was determined (h orth);
#          if this was 30 the gene was considered to be a class B gene.
# Class A: Class A genes were defined as a subset of class B genes with no recipient matches with bitscore 100 and
#          no members of their respective ortholog group with recipient matches with bitscore 100.


# get the directory our bacterial genome GFF annotation files are in
bac_gff_directory <-
  here("data","01_psam_intergenic_HGT_index","psam_bac_genomes")


# list the gff files in each directory as a variable
bac_gff_files <-
  list.files(path = bac_gff_directory, full.names = T, pattern = "\\.gff")


# Here we use fread, lapply, and bind_rows to read in all of our GFF files and bind them by row (i.e., concatenating them) into a single data frame.
bac_gff <-
  bind_rows(lapply(bac_gff_files, fread, skip=1, header=F, 
col.names = c("chrm_bac","source","feature_type","feature_start","feature_end","score","strand","phase","attributes"))) %>%
  filter(feature_type == "CDS") %>% 
  mutate(attributes = str_remove(attributes, "ID\\=fig\\|203682\\.28\\.peg\\.|ID\\=fig\\|41295\\.5\\.peg\\.|ID\\=fig\\|89373\\.5\\.peg\\.|ID\\=fig\\|1813606\\.4\\.peg\\.")) %>% 
  separate(col = attributes, into = c("bac_gen_peg", "bac_gen_gene_annotation"), sep = ";") %>% 
  mutate(bac_gen_gene_annotation = str_remove(bac_gen_gene_annotation, "Name\\=")) %>% 
  mutate(length_bac = paste(feature_start,"-",feature_end, "(", strand, ")", sep = "")) %>%
  select(-c("feature_type","feature_start","feature_end","score","strand")) %>% 
  mutate(chrm_bac = str_remove(chrm_bac, pattern = "^00000")) 

# read in our orthogroup information
orthogroups_total <- 
  read_tsv(here::here('data','02_hgt_index','orthofinder_psammoneis_wade','OrthoFinder','Results_May12','Orthogroups','Orthogroups.tsv')) %>% 
  as.data.frame() %>% 
  clean_names() %>% 
  mutate(proteomes_per_OG = apply(across(2:13), 1, count_nonNA_func)) %>% 
  mutate(grouping = 
             case_when(
              is.na(psammoneis_japonica) &
                is.na(nitzschia_nitz4) &
                is.na(pseudonitzschia_multistriata) &
                is.na(fistulifera_solaris) &
                is.na(pseudonitzschia_multiseries) &
                is.na(fragilariopsis_cylindrus) &
                is.na(phaeodactylum_tricornutum) &
                is.na(psammoneis_japonica) &
                is.na(seminavis_robusta) &
                is.na(ectocarpus_siliculosus) &
                is.na(nannochloropsis_gaditana) ~ "centric",
              is.na(psammoneis_japonica) &
                is.na(nitzschia_nitz4) &
                is.na(pseudonitzschia_multistriata) &
                is.na(pseudonitzschia_multiseries) &
                is.na(fragilariopsis_cylindrus) &
                is.na(cyclotella_cryptica) &
                is.na(thalassiosira_pseudonana) &
                is.na(ectocarpus_siliculosus) &
                is.na(nannochloropsis_gaditana) ~ "raphid_clade_1",
              is.na(psammoneis_japonica) &
                is.na(fistulifera_solaris) &
                is.na(phaeodactylum_tricornutum) &
                is.na(seminavis_robusta) &
                is.na(cyclotella_cryptica) &
                is.na(thalassiosira_pseudonana) &
                is.na(ectocarpus_siliculosus) &
                is.na(nannochloropsis_gaditana) ~ "raphid_clade_2",
              is.na(psammoneis_japonica) &
                is.na(cyclotella_cryptica) &
                is.na(thalassiosira_pseudonana) &
                is.na(ectocarpus_siliculosus) &
                is.na(nannochloropsis_gaditana) ~ "deep_raphid",
                is.na(cyclotella_cryptica) &
                is.na(thalassiosira_pseudonana) &
                is.na(ectocarpus_siliculosus) &
                is.na(nannochloropsis_gaditana) ~ "pennate",
               is.na(ectocarpus_siliculosus) & 
                 is.na(nannochloropsis_gaditana) ~ "diatom"
             )
  ) %>% 
  replace_na(list(grouping = "all")) %>% 
  filter(!proteomes_per_OG == 1) %>% # This filters out terminal orthogroups (species specific orthogroups) so that we can look at non-terminal (at the nodes)
  mutate(across(
    .cols = 2:13,
    .fns = ~ str_count(.x, ',')+1, 
    .names = "{col}_total_genes")) %>% # use across to calculate # of genes per taxon in each OG and write the output as new columns
  mutate(across(
    .cols = ends_with("_total_genes"), 
    .fns = ~ ifelse(is.na(.x), 0, .x)))


  
orthogroups_cafe <- 
  read_tsv(here::here('data','08_gene_families','CAFE','Orthogroups.GeneCount.CAFE.txt')) %>% 
  as.data.frame() %>% 
  clean_names()  %>% 
  
orthogroups_cafe_psammoneis <- 
  read_tsv(here::here('data','08_gene_families','CAFE','Orthogroups.GeneCount.CAFE.txt')) %>% 
  as.data.frame() %>% 
  clean_names() %>% 
  filter(psammoneis != 0)
  
orthogroups_psam <- 
  read_tsv(here::here('data','02_hgt_index','orthofinder_psammoneis_wade','OrthoFinder','Results_May12','Orthogroups','Orthogroups_Psam_filtered.tsv')) %>% 
  as.data.frame() %>% 
  clean_names() %>% 
  mutate(across(
    .cols = 2:13, 
    ~ str_count(.x, ',')+1, 
    .names = "{col}_total_genes")) %>% # use across to calculate # of genes per taxon in each OG and write the output as new columns
  mutate(across(
    .cols =ends_with("_total_genes"), 
    ~ ifelse(is.na(.x), 0, .x))) # go through all of the newly created columns and convert NAs to zeroes

# check that we have 15170 genes in total
orthogroups_psam %>% 
  summarise(psam_gene_total = sum(psammoneis_japonica_total_genes)) 

# read in unnassigned genes from orthofidner
orthogroups_unassigned <- 
    read_tsv(here::here('data','02_hgt_index','orthofinder_psammoneis_wade','OrthoFinder','Results_May12','Orthogroups','Orthogroups_UnassignedGenes.tsv')) %>% 
  as.data.frame() %>% 
  clean_names()

# read in UniProt information for each sseqid in our blast databases
uniprot_reviewed <-
  read_tsv(file = here("data","uniprot-reviewed_yes.tsv"), col_names = T) %>%
  as.data.frame() %>%
  clean_names()

# assign the output directories as variables
donor_directory <-
  here("data","02_hgt_index","HGT_index_output","HGT_index_output_donor")
recipient_directory <-
  here("data","02_hgt_index","HGT_index_output","HGT_index_output_recipient")

# list the files in each directory as a variable
donors <-
  list.files(path = donor_directory, full.names = T)
recipients <-
  list.files(path = recipient_directory, full.names = T)

# Here we use fread, lapply, and bind_rows to read in all of our output files and bind them by row (i.e., concatenating them) into a single dataframe for donor and recipient results, respectively.
donor_files <-
  bind_rows(lapply(donors, fread)) %>%
  group_by(qseqid) %>% # group our data by sqeqid
  top_n(1, bitscore_d) %>% # select the top blast search result by bitscore for each seq_id
  slice(1) %>%
  mutate(orthogroup = str_remove(orthogroup, ".fa.donor.dmndBlastOut.tsv")) %>% 
  ungroup()

recipient_files <-
  bind_rows(lapply(recipients, fread)) %>%
  group_by(qseqid) %>% # group our data by sqeqid
  top_n(1, bitscore_r) %>% # select the top blast search result by bitscore for each seq_id
  slice(1) %>%
  mutate(orthogroup = str_remove(orthogroup, ".fa.recipient.dmndBlastOut.tsv")) %>% 
  ungroup()

# We will use this to examine what percentage of our class A hits are only to our 0F-3F bacterial genomes
# and which are also hitting to uniprot hits
donor_files_all <-
  bind_rows(lapply(donors, fread)) %>%
  group_by(qseqid) %>% # group our data by sqeqid
  mutate(orthogroup = str_remove(orthogroup, ".fa.donor.dmndBlastOut.tsv")) %>% 
  ungroup()


# calculate HGT index per orthogroup (h orth)
# This needs to be done first because R gets grumpy with you when trying this and everything else in a single, piped command.
d_r_hOrth <-
  full_join(donor_files, recipient_files, by = c("qseqid", "orthogroup")) %>% # join donor and recipient output by qseqid and orthogroup columns
  replace_na(list(
    bitscore_d = 0,
    bitscore_r = 0
  )) %>% # here we replace all null values in recipient and donor bitscores with zeros. This is necessary to assign our 'strong' domain classes.
  mutate(h = bitscore_d - bitscore_r) %>% # calculate HGT index (h)
  group_by(orthogroup) %>% # necessary to calculate h_orth
  dplyr::summarise(dplyr::across(
    .cols = dplyr::starts_with("h"),
    ~ mean(.x, na.rm = TRUE),
    .names = "{.col}_orth"
  )) %>% # Calculate h_orth for each orthogroup (mean h per orthogroup)
  naniar::replace_with_na(., replace = list(h_orth = "NaN")) # here we replace all NaN values introduced by h_orth calculation with actual NA values.

# Here we perform our final joins and calculate and assign the crisp GT index classes.
hgt_index_output <-
  full_join(donor_files, recipient_files, by = c("qseqid", "orthogroup")) %>%
  replace_na(list(
    bitscore_d = 0,
    bitscore_r = 0
  )) %>% # here we replace all null values in recipient and donor bitscores with zeros. This is necessary to assign our 'strong' domain classes.
  mutate(h = bitscore_d - bitscore_r) %>% # calculate HGT index (h)
  full_join(., d_r_hOrth, by = "orthogroup") %>%
  mutate(
    domain_assignment = # create domain assignment column
      case_when(
        # strong prokaryote
        bitscore_r == 0 & h >= 30 ~ "prokaryote_strong",
        # strong eukaryote
        bitscore_d == 0 & h <= 0 ~ "eukaryote_strong",
        # prokaryote
        h >= 30 ~ "prokaryote",
        # eukaryote
        h <= 0 ~ "eukaryote",
        # indeterminate
        between(h, 0, 30) ~ "indeterminate"
      )
  )

hgt_index_output_orth <-
  left_join(hgt_index_output,
    (hgt_index_output %>%
      group_by(orthogroup) %>%
      summarise(max_bitscore_r_orth = max(bitscore_r, na.rm = TRUE))),
    by = "orthogroup"
  ) %>% # here we join a summary df of the max value in bitscroe_r for each orthogroup to the main dataframe, this allows for calculation of class A hgt classes
  mutate(
    hgt_class = # create our hgt class column, we need to assign classes from least to most inclusive because case_when will overwrite previous assignments that meet the criteria of the next class.
      case_when(
        # Class A: Class A genes were defined as a subset of class B genes with no recipient matches with bitscore 100 and
        #          no members of their respective ortholog group with recipient matches with bitscore 100.
        h_orth >= 30 & bitscore_r < 100 & max_bitscore_r_orth < 100 & (domain_assignment == "prokaryote" |domain_assignment == "prokaryote_strong") ~ "A",
        # Class B: For each class C gene, the average h value of the members of its ortholog group was determined (h orth);
        #          if this was 30 the gene was considered to be a class B gene.
        h_orth >= 30 & (domain_assignment == "prokaryote" |domain_assignment == "prokaryote_strong")  ~ "B",
        # Class C: h >= 30 & best donor bitscore >= 100
        h >= 30 & bitscore_d >= 100 & (domain_assignment == "prokaryote" |domain_assignment == "prokaryote_strong")  ~ "C"
      )
  ) %>%
   separate(sseqid_r, sep = "\\|", into = c("uniprot_prefix_r", "uniprot_id_r", "product_r"), remove = T) %>% # sseqid_r split
  separate(sseqid_d, sep = "\\|", into = c("uniprot_prefix_d", "uniprot_id_d", "product_d"), remove = F) %>% # uniprot hits
  separate(sseqid_d, sep = "\\.", into = c("chrm_bac", "bac_gen_peg"), remove = F) %>% # bacterial genome hits
  left_join(., uniprot_reviewed, by = c("uniprot_id_r" = "entry")) %>% # Join Uniprot info to recipient hits
  select(-c(starts_with("uniprot_prefix"))) %>%
  dplyr::rename(
    entry_name_uniprot_r = entry_name, gene_ontology_biological_process_uniprot_r = gene_ontology_biological_process, gene_ontology_molecular_function_uniprot_r = gene_ontology_molecular_function, gene_names_uniprot_r = gene_names,
    organism_uniprot_r = organism, taxonomic_lineage_superkingdom_r = taxonomic_lineage_superkingdom
  ) %>%
  left_join(., uniprot_reviewed, by = c("uniprot_id_d" = "entry")) %>% # Join Uniprot info to donor hits
  dplyr::rename(
    entry_name_uniprot_d = entry_name, gene_ontology_biological_process_uniprot_d = gene_ontology_biological_process, gene_ontology_molecular_function_uniprot_d = gene_ontology_molecular_function, gene_names_uniprot_d = gene_names,
    organism_uniprot_d = organism, taxonomic_lineage_superkingdom_d = taxonomic_lineage_superkingdom
  ) %>%
  left_join(., bac_gff, by = c("chrm_bac","bac_gen_peg")) %>% # join our gff info
  as.data.frame() %>% 
  mutate(taxonomic_lineage_superkingdom_d = ifelse(is.na(taxonomic_lineage_superkingdom_d), "Bacteria", taxonomic_lineage_superkingdom_d)) %>% # replace NAs in superkingdom column with bacteria for our bacterial genome hits
  mutate(
      organism_uniprot_d = # assign taxonomy of our bacterial genomes tp organism_uniprot_d column
        case_when(
          # strong prokaryote
          chrm_bac == "0F" ~ "Phycisphaerae sp.",
          chrm_bac == "1F" ~ "Rhodovibrionaceae sp.",
          chrm_bac == "2F" ~ "Ekhidna sp.",
          chrm_bac == "3F" ~ "Balneola sp.",
        )
    ) %>% 
    mutate(bac_genome =
           case_when(
             str_detect(sseqid_d, '0F\\.[:digit:]') ~ "0F",
             str_detect(sseqid_d, '1F\\.[:digit:]') ~ "1F",
             str_detect(sseqid_d, '2F\\.[:digit:]') ~ "2F",
             str_detect(sseqid_d, '3F\\.[:digit:]') ~ "3F"
           ))

# prokaryotic subset of hits
hgt_index_output_orth_prok_psam <-
  hgt_index_output_orth %>%
  filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
  filter(grepl("^Psammoneisjap", qseqid))

## We will use this to examine what percentage of our class A hits are only to our 0F-3F bacterial genomes
# and which are also hitting to uniprot hits
bac_gen_unqiue_hits_qseqid <- 
  left_join(
  (donor_files_all %>% filter(grepl("^Psammoneisjap", qseqid))), 
   (hgt_index_output_orth_prok_psam %>% filter(hgt_class == 'A')), 
          by = c("qseqid", "sseqid_d", "orthogroup", "pident_d","length_d","mismatch_d","gapopen_d","qstart_d","qend_d","sstart_d","send_d","evalue_d","bitscore_d")) %>% 
  select(-c(ends_with("_r"))) %>% 
  group_by(qseqid) %>%
  filter(any(!is.na(h))) %>% 
  filter(any(bac_genome %in% c("0F","1F","2F","3F"))) %>% 
  select(orthogroup, qseqid, sseqid_d, h, h_orth, hgt_class, chrm_bac) %>% 
  mutate(total_hits = n()) %>% 
  # ungroup() %>% 
  # group_by(orthogroup) %>% 
  mutate(qseqid_distinct = n_distinct(qseqid)) %>% 
  ungroup() %>% 
  group_by(qseqid) %>% 
  mutate(genome_pct = sum(str_count(sseqid_d, '[:digit:]+F'))/n()) %>% 
  arrange(orthogroup, desc(h)) %>% 
  ungroup()

rm(uniprot_reviewed, d_r_hOrth, donor_files, recipient_files, recipients, donors, donor_directory, recipient_directory)
```

```{r}
#| label: find h-cutoff genic
#| eval: true
#| echo: false
#| warning: false
#| error: false
#| output: false
hgt_index_output_orth %>%
  ggplot(aes(x = h)) +
  stat_ecdf(geom = "point") +
  scale_x_continuous(breaks = c(seq(-200, 300, by = 50), 30), limits = c(-200,300)) +
  xlab("HGT index (<i>h</i>)") +
  ylab("Empirical cumulative distribution of <i>h</i>") +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    # panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "none",
    legend.title = element_blank()
  ) +
  geom_vline(xintercept = 30, color = "black", linetype = 3)
```

```{r}
#| label: Genic data summaries and exploration
#| eval: false
#| echo: true
#| warning: false
#| error: false
#| output: false


# check out the results for the orthogroups for unassigned genes
hgt_index_output_orth_prok_psam_unassigned <- 
  hgt_index_output_orth_prok_psam %>% 
  filter(orthogroup %in% orthogroups_unassigned$orthogroup)


bac_gen_unqiue_hits_OG <- 
  left_join(
  (donor_files_all %>% filter(grepl("^Psammoneisjap", qseqid))), 
   (hgt_index_output_orth_prok_psam %>% filter(hgt_class == 'A')), 
          by = c("qseqid", "sseqid_d", "orthogroup", "pident_d","length_d","mismatch_d","gapopen_d","qstart_d","qend_d","sstart_d","send_d","evalue_d","bitscore_d")) %>% 
  select(-c(ends_with("_r"))) %>% 
  group_by(orthogroup) %>%
  filter(any(!is.na(h))) %>% 
  filter(any(bac_genome %in% c("0F","1F","2F","3F"))) %>% 
  select(orthogroup, qseqid, sseqid_d, h, h_orth, hgt_class, chrm_bac) %>% 
  mutate(qseqid_distinct = n_distinct(qseqid)) %>% 
  mutate(total_hits = n()) %>% 
  ungroup() %>% 
  group_by(orthogroup) %>% 
  mutate(genome_pct = sum(str_count(sseqid_d, '[:digit:]+F'))/n()) %>% 
  arrange(orthogroup, desc(h)) %>% 
  ungroup()

# plotting total hits per grouping and percentage of hits to bacterial genomes, faceted by group with boxplot ggside
rbind(
(bac_gen_unqiue_hits_qseqid %>% 
   select(qseqid, total_hits, genome_pct) %>% 
   unique() %>% 
   select(-qseqid) %>% 
   add_column(grouping = "qseqid")
    ), 
(bac_gen_unqiue_hits_OG %>% 
   select(orthogroup, total_hits, genome_pct) %>% 
   unique() %>% 
   select(-orthogroup) %>% 
   add_column(grouping = "OG")
    ) 
) %>% 
  ggplot(aes(x = total_hits, y = genome_pct)) +
  ggpointdensity::geom_pointdensity() +
  viridis::scale_color_viridis(begin = 0, end = 0.75) +
  labs(y="Percentage of hits to bacterial genomes 0F - 3F", 
       x="Total no. of hits per group") +
  facet_wrap(~ grouping, ncol = 1, scales = "free", strip.position = "top") +
  theme(
    # axis.text.x = element_text(angle = 0, hjust = .5, color = "#000000"),
    # axis.text.y = element_text(color = "#000000", size = 10),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    # panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = c(.75, .2),
    legend.title = element_blank(),
    legend.background = element_blank()
  ) +
  geom_ysideboxplot(aes(x = genome_pct), orientation = "x", outlier.color = "gray") +
  geom_xsideboxplot(aes(x = total_hits), orientation = "y", outlier.color = "gray")


# subset to just hits to our bacteria genomes
hgt_index_output_orth_prok_psam %>% 
  filter(str_detect(sseqid_d, '[:digit:]F\\.[:digit:]')) %>% 
  select(orthogroup, bac_genome, bac_gen_peg) %>% 
  table() %>% 
  data.frame() %>% 
  arrange(desc(Freq)) %>% 
  filter(Freq >= 1)

# Just the bac_genome and peg number frequencies
hgt_index_output_orth_prok_psam %>% 
  filter(str_detect(sseqid_d, '[:digit:]F\\.[:digit:]')) %>% 
  select(bac_genome, bac_gen_peg, bac_gen_gene_annotation) %>% 
  table() %>% 
  data.frame() %>% 
  arrange(desc(bac_genome), desc(bac_gen_peg)) %>% 
  filter(Freq >= 3)

# Here we will filter to see if any of the sseqid_d values match the genome.PEG IDs fthat Matt recovered
filter(hgt_index_output_orth_prok_psam, sseqid_d %in% c("0F.2931","0F.660","0F.766","0F.3987","0F.4510","3F.313","0F.1819")) %>% 
  select(where(~ !(all(is.na(.)) | all(. == "")))) # Drop empty columns

# Psammoneis subset of hits by domain assignment
# Number of unique orthogroups per domain assignment for psammoneis
hgt_index_output_orth %>%
  filter(grepl("^Psammoneisjap", qseqid)) %>%
  group_by(domain_assignment) %>% 
  summarise(cnt = dplyr::n(), unique_orthogroups = length(unique(orthogroup)))

# Class assignments for all hits
 hgt_index_output_orth %>%
  group_by(domain_assignment) %>%
  summarise(cnt = dplyr::n()) %>%
  mutate(freq = round(cnt / sum(cnt), 3)) %>%
  dplyr::arrange(desc(freq))

#domain assignments by origin of hit
hgt_index_output_orth %>%
  filter(grepl("^Psammoneisjap", qseqid)) %>%
  select(domain_assignment, bac_genome) %>%
    table(useNA = "ifany") %>% 
   as_tibble() %>% 
   dplyr::mutate(bac_genome = replace_na(bac_genome, "uniprot"))


 # Domain assignment summaries
# Domain assignments for prokaryotic hits for psammoneis
    hgt_index_output_orth %>%
  filter(grepl("^Psammoneisjap", qseqid)) %>%
  filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
  select(taxonomic_lineage_superkingdom_d) %>%
    table(useNA = "ifany") 
    
# Domain assignments for prokaryotic hits by origin of hit
hgt_index_output_orth %>%
  filter(grepl("^Psammoneisjap", qseqid)) %>%
  filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
  select(taxonomic_lineage_superkingdom_d, bac_genome) %>%
    table(useNA = "ifany") %>% 
   as_tibble() %>% 
   dplyr::mutate(bac_genome = replace_na(bac_genome, "uniprot"))

# Class assignment summaries
# class assignments
   hgt_index_output_orth %>%
  filter(grepl("^Psammoneisjap", qseqid)) %>%
  filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
  select(hgt_class) %>%
    table(useNA = "ifany") 

   # bacteria genome subset
  hgt_index_output_orth %>%
  filter(grepl("^Psammoneisjap", qseqid)) %>%
  filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
  filter(bac_genome %in% c("0F","1F","2F","3F")) %>%
  select(where(~ !(all(is.na(.)) | all(. == "")))) %>% # Drop empty columns
  select(hgt_class) %>%
    table(useNA = "ifany") 
   
# class assignments by hit origin
   hgt_index_output_orth %>%
  filter(grepl("^Psammoneisjap", qseqid)) %>%
  filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
  select(hgt_class, bac_genome) %>%
    table(useNA = "ifany") %>% 
   as_tibble() %>% 
   dplyr::mutate(bac_genome = replace_na(bac_genome, "uniprot"))

# class hits by domain archaea      
  hgt_index_output_orth %>%
  filter(grepl("^Psammoneisjap", qseqid)) %>%
  filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
    filter(taxonomic_lineage_superkingdom_d == "Archaea") %>%
  select(hgt_class) %>%
    table(useNA = "ifany") 

# class hits by domain archaea by origin of hit
# obviously it will only return uniprot as a value because all of our genomes are from bacteria
  hgt_index_output_orth %>%
  filter(grepl("^Psammoneisjap", qseqid)) %>%
  filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
    filter(taxonomic_lineage_superkingdom_d == "Archaea") %>%
  select(hgt_class, bac_genome) %>%
    table(useNA = "ifany") %>% 
   as_tibble() %>% 
   dplyr::mutate(bac_genome = replace_na(bac_genome, "uniprot"))

# class hits by domain bacteria      
  hgt_index_output_orth %>%
  filter(grepl("^Psammoneisjap", qseqid)) %>%
  filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
    filter(taxonomic_lineage_superkingdom_d == "Bacteria") %>%
  select(hgt_class) %>%
    table(useNA = "ifany") 

# class hits by domain bacteria by origin of hit
  hgt_index_output_orth %>%
  filter(grepl("^Psammoneisjap", qseqid)) %>%
  filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
    filter(taxonomic_lineage_superkingdom_d == "Bacteria") %>%
   dplyr::mutate(bac_genome = replace_na(bac_genome, "uniprot")) %>% 
  select(hgt_class) %>%
    table(useNA = "ifany") %>% 
   as_tibble()

    hgt_index_output_orth %>%
  filter(grepl("^Psammoneisjap", qseqid)) %>%
  filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
    filter(taxonomic_lineage_superkingdom_d == "Bacteria") %>%
      filter(bac_genome %in% c("0F","1F","2F","3F")) %>%
  select(hgt_class) %>%
    table(useNA = "ifany") %>% 
   as_tibble() 

# Missing data visualization
naniar::vis_miss(orthogroups)
naniar::gg_miss_upset(orthogroups, nsets = 12, order.by = "freq")

# columns that would need to be NA for Psammoneis only filtering
orthogroups_psam_cols <-
  colnames(orthogroups_psam %>% select(-psammoneis_japonica, -orthogroup))

orthogroups_psam_CAFE_cols <-
  colnames(orthogroups_cafe %>% select(-c(psammoneis, orthogroup, desc)))

# Orthogroups with multiple diatoms
orthogroups_diatom_only <-
  orthogroups_psam %>% 
  select(-nannochloropsis_gaditana, -ectocarpus_siliculosus) %>% 
      filter(!if_all(colnames(orthogroups_psam %>% select(-psammoneis_japonica, -orthogroup, -nannochloropsis_gaditana, -ectocarpus_siliculosus)), ~ is.na(.))) 

# Filtering diatom data to include only Phaeo and Seminavis
orthogroups_phaeo_seminavis <-
  orthogroups_diatom_only %>% 
  filter(!if_all(c('phaeodactylum_tricornutum', 'seminavis_robusta'), ~ is.na(.))) 

# subset orthogroups where psammoneis is the only gene donor
psam_only_orths <-  
  orthogroups_psam%>% 
  filter(if_all(all_of(orthogroups_psam_cols), ~ is.na(.))) %>% 
  select(orthogroup, psammoneis_japonica) %>% 
  mutate(gene_count = str_count(psammoneis_japonica, ',')+1)

psam_only_orths_CAFE <-  
  orthogroups_cafe %>% 
  filter(if_all(all_of(orthogroups_psam_CAFE_cols), ~ . == 0)) %>% 
  select(orthogroup, psammoneis) %>% 
  mutate(gene_count = str_count(psammoneis, ',')+1)

# see which of our psam only OGs are in our HGT output
left_join(psam_only_orths, hgt_index_output_orth_prok_psam, by = "orthogroup")%>%
      filter(bac_genome %in% c("0F","1F","2F","3F")) %>%
  select(hgt_class) %>%
    table(useNA = "ifany") %>% 
   as_tibble() 

# let's see which singleton OGs are in Psammoneis
# end up with an empty DF, so all of these are not singletons
psam_only_orths %>% 
  filter(!str_detect(psammoneis_japonica, ','))

# Let's try the reverse and filter out the HGT hits and then visualize the missingness of our data
orthogroups_hgt_hits <-
  orthogroups_psam %>% 
  filter(orthogroup %in% hgt_index_output_orth$orthogroup)

# Missing Data visualization
naniar::vis_miss(orthogroups_hgt_hits)
naniar::gg_miss_upset(orthogroups_hgt_hits, nsets = 12, order.by = "freq")
naniar::gg_miss_var(orthogroups_hgt_hits)

# let's look at qseqs that are in psammoneis only orthogroups
# 196 hits...82 uniques, like in the upset plot
hgt_index_output_orth_psam_only <-
  hgt_index_output_orth %>% 
  filter(orthogroup %in% psam_only_orths$orthogroup) 

# Let's also check on what HGT classes are from orthogroups with Phaeodactylum and Seminavis
hgt_index_output_orth_pheo_seminavis <-
  hgt_index_output_orth %>% 
  filter(orthogroup %in% orthogroups_phaeo_seminavis$orthogroup) 

# Of the Psammoneis prokaryotic HGT hits, figure out the ratio of Psammoneis only OGs to multi-diatom OGs
# Psammoneis_only
left_join(hgt_index_output_orth_prok_psam, orthogroups, by = "orthogroup") %>% 
  filter(orthogroup %in% psam_only_orths$orthogroup)
  

# Multi-diatom
left_join(hgt_index_output_orth_prok_psam, orthogroups, by = "orthogroup") %>% 
  filter(orthogroup %in% orthogroups_diatom_only$orthogroup)

# Then we will figure out, of the HGT hits to our Bacterial genomes, which are Psammoneis-only and which are multi-diatom 
# Psammoneis_only
left_join(hgt_index_output_orth_prok_psam, orthogroups, by = "orthogroup") %>% 
   filter(bac_genome %in% c("0F","1F","2F","3F")) %>% 
  filter(orthogroup %in% psam_only_orths$orthogroup) 

# Multi-diatom
left_join(hgt_index_output_orth_prok_psam, orthogroups, by = "orthogroup") %>% 
   filter(bac_genome %in% c("0F","1F","2F","3F")) %>% 
  filter(orthogroup %in% orthogroups_diatom_only$orthogroup)
```

Finally, we will use NCBI BlastP to check our class A sequences.  
First, we use the SLURM script 'submit_01_download_ncbi_nr_db_c1432.slurm' to download the most recent NR blast database and taxonomy database. 

Next, we export our class A Psammoneis protein sequences into a FASTA file. 


```{r}
#| label: export class A seqs for ncbi blastp analysis
#| eval: true
#| echo: false
#| warning: false
#| error: false
#| output: false



# We will import Psammoneis CDS sequences from orthofinder and filter them to only the sequences of Class A hits to our bacterial genome
psam_CDS <-
  bioseq::read_fasta(file = here("data","02_hgt_index","orthofinder_psammoneis_wade","Psammoneis_japonica.faa"), type = "AA")

# We'll then export those and NCBI blastp them to check that the top hit is prokaryotic
# Psammoneis_only
psam_qseqs_classA <-
  hgt_index_output_orth_prok_psam %>% 
  filter(hgt_class == "A") %>% 
  select(qseqid) %>% 
  tibble()

# now we filter the Psammoneis CDS seqs with the class A qseqids
psam_qseqs_classA_seqs <-
  tibble(label = names(psam_CDS), sequence = psam_CDS) %>% 
  filter(label %in% psam_qseqs_classA$qseqid) %>% 
  mutate(label = paste(">", label, sep = ""))


# split up the psam_CDS_classA_NONbacgen file into chunks so that ncbi BlastP doesn't throw a fit
write.table(psam_qseqs_classA_seqs, 
                 file = here("data","02_hgt_index","orthofinder_psammoneis_wade","Psammoneis_japonica_classA_seqs.faa"), 
                          sep = "\n", 
                          quote = FALSE, 
                          row.names = FALSE, 
                          col.names = FALSE
                          )
```

After we do that, we use them as the query sequences for a BLASTP search using the following command:
```
blastn -db nt -num_threads 24 -max_target_seqs 10 -outfmt '6 qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore sacc staxids sscinames' -query Psammoneis_japonica_classA_seqs.faa
```

One important step to note, to have blast output with staxids and sscinames, we need the taxdb files from NCBI. Those files must be located in the directory from which you initiate the blast command, NOT the directory the blast DB is located in.

After NCBI BlastP runs, we also download the prot.accession2taxid files from [NCBI](https://ftp.ncbi.nih.gov/pub/taxonomy/accession2taxid/)
We extract the contents of the .gz files and join them using the following one-liner in the terminal. First, we concatenate all of the extracted prot.accession2taxid.FULL tsv files, then pipe that to grep, where we use the -v option to get the inverse match to taxid. This will remove all of the header lines from our output. Finally the output from grep is directed to our final file, prot.accession2taxid.FULL.2022-10-09.tsv.

```
cat prot.accession2taxid.FULL.* | grep -v 'taxid' > prot.accession2taxid.FULL.2022-10-09.tsv
```

```{r}
#| label: taxid2accession file to parquet
#| eval: false
#| echo: false
#| warning: false
#| error: false
#| output: false
#| 
# One thing we want to do to make things easier is use the prot.accessions2taxid file for assigning sequence origin to our results. Unfortuantely, the full size of this database is >86gb
# We have two options for dealing with this: 
# 1) use an sqlite database
# 2) create a parquet file directory to use Arrow to parse our data.

# I opted for the parquet database because operations are a bit faster and the size of the parquet files is much smaller at ~30gb. 

# Here, I have set up a workflow to convert our tsv file into a parquet db 
tsv_file_dir <- "~/local/ncbi/prot.accession2taxid.FULL_2022-10-09_tsvs"
dest <- "~/local/ncbi/prot.accession2taxid.FULL_2022-10-09_parquet/" 

sch = arrow::schema(accession.ver = string(),
                    taxid = numeric())

csv_stream <- open_dataset(tsv_file_dir, format = "tsv", 
                           schema = sch, skip_rows = 1)

write_dataset(csv_stream, dest, format = "parquet", 
              max_rows_per_file=1000000L,
              hive_style = TRUE,
              existing_data_behavior = "overwrite")

# Convert our gff tsv file to a parquet file
tsv_file <- here("data","01_psam_intergenic_HGT_index","psam_genome","psamm_final.round4.all.gff_filtered.tsv")
dest <- here("data","01_psam_intergenic_HGT_index","psam_genome","psamm_final.round4.all.gff_parquet")

sch <-  arrow::schema(seqname = string(),
                      source = string(),
                      feature = string(),
                      start = numeric(),
                      end = numeric(),
                      score = string(),
                      strand = string(),
                      frame = string(),
                      attribute = string())

tsv_stream <- open_dataset(tsv_file, format = "tsv", 
                           schema = sch, skip_rows = 0)

write_dataset(tsv_stream, dest, format = "parquet", 
              max_rows_per_file=1000000L,
              hive_style = TRUE,
              existing_data_behavior = "overwrite")
```


```{r}
#| label: import and explore ncbi blastp results
#| eval: true
#| echo: false
#| warning: false
#| error: false
#| output: false


# open our arrow datasets
prot.accession2taxid <- 
  open_dataset("/Users/cbg/local/ncbi/prot.accession2taxid.FULL_2022-10-09_parquet")

psam_gff <- 
  open_dataset(here("data","01_psam_intergenic_HGT_index","psam_genome","psamm_final.round4.all.gff_parquet"))


# filter psam_gff by mRNA to get coords for our pdammoneis proteome (that we used in orthofinder)
psam_gff_proteome <- 
  psam_gff %>% 
  filter(feature == "mRNA") %>% 
  collect()

# filter the above
psam_gff_filtered <- 
psam_gff_proteome %>% 
  mutate(id = str_remove_all(attribute, paste(c("^ID=", "\\;.+$","\\:.+$"), collapse = "|"))) %>%
  select(-attribute) %>% 
  filter(id %in% # filter our gff file by ids from our HGT index output
           (hgt_index_output_orth_prok_psam %>% # remove psamjap from qseqids for filtering of gff parquet
            mutate(id = str_remove(qseqid, "^Psammoneisjap_")) %>% 
            select(id))$id) %>% 
  mutate(qseqid = paste("Psammoneisjap_", id, collapse = ""))

# Read in the NCBI BlastP results
blastp_results <-
  read_tsv(
    file = here("data","02_hgt_index","ncbi_blastp_output","Psammoneis_japonica_classA_seqs_blastp_out.tsv"),
    col_names = c("qseqid", "sseqid", "pident", "length", "mismatch", "gapopen", "qstart", "qend", "sstart", "send", "evalue", "bitscore", "sacc", "staxids", "sscinames"), 
    progress = T) %>% 
  mutate(acc.ver = str_remove_all(sseqid, paste(c("^[:alpha:]+\\|+","\\|$","\\|[:alpha:]$"), collapse="|"))) 

# read in taxid categories file
taxid_categories <-
  read_tsv(
    file = "/Users/cbg/local/ncbi/categories.dmp",
    col_names = c("category", "taxid_species_level", "taxid"), 
    progress = T) %>% 
  mutate(category_expanded = case_when(
  category == "A" ~ "Archaea",
  category == "B" ~ "Bacteria",
  category == "E" ~ "Eukaryota",
  category == "V" ~ "Viruses and Viroids",
  category == "U" ~ "Unclassified",
  category == "O" ~ "Other"
  ))


# join our datasets
blastp_results <- 
  left_join( # join our blastp results with our prot.accession2taxid data from the parquet files
    blastp_results,
    (prot.accession2taxid %>% 
  filter(accession.ver %in% blastp_results$acc.ver) %>% 
  collect()),
  by = c("acc.ver" = "accession.ver")
  ) %>% 
  left_join(., taxid_categories, by = "taxid") # Join our taxid_categories file to the above

  
blastp_results_joined <- 
  left_join(blastp_results,
            (blastp_results %>% 
               group_by(qseqid) %>% 
               select(qseqid, category_expanded) %>% 
               table() %>% 
               as.data.frame() %>% 
               pivot_wider(names_from = category_expanded, values_from = Freq) %>% 
               mutate( # now we make a column to calculate the total hits to each qseqid
                 total_hits = rowSums(select(., contains(c("Arc","Bac","Euk"))))
                 )),
            by = "qseqid")

# Next let's make a stacked bar plot to clarify the composition of our top 10 blast hits and the top hit that each qseqid hit to
# New labels for facets
# New facet label names for supp variable
supp.labs <- c("Top hit to Archaea", "Top hit to Bacteria", "Top hit to Eukaryota")
names(supp.labs) <- c("Archaea", "Bacteria", "Eukaryota")

  # plot it but with just our 106 class A proteins that hit to psammoneis! 
    left_join(
      (hgt_index_output_orth %>%
         filter(grepl("^Psammoneisjap", qseqid)) %>%
         filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
         filter(bac_genome %in% c("0F","1F","2F","3F")) %>% 
         filter(hgt_class == "A")),
      blastp_results_joined,
      by = 'qseqid') %>% 
    group_by(qseqid) %>% 
    top_n(10, bitscore) %>% 
    slice_head(n = 10) %>% 
    ungroup() %>% 
    select(qseqid, category_expanded) %>% 
    table() %>% 
    as.data.frame() %>% 
    pivot_wider(names_from = category_expanded, values_from = Freq) %>% 
    mutate( # now we make a column to calculate the total hits to each qseqid
    total_hits = rowSums(select(., contains(c("Arc","Bac","Euk"))))
    # perc_arc = Archaea/total_hits,
    # perc_bac = Bacteria/total_hits,
    # perc_euk = Eukaryota/total_hits
    ) %>% 
    pivot_longer(
      cols = starts_with(c("Arc","Bac","Euk")),
      names_to = "groups",
      values_to = "values",
      values_drop_na = TRUE
      ) %>% 
    # pivot_longer(
    #   cols = starts_with("perc"),
    #   names_to = "perc_groups",
    #   values_to = "perc_values",
    #   values_drop_na = TRUE
    #   ) %>% 
    left_join(
      ., 
      (blastp_results_joined %>% 
        group_by(qseqid) %>% # group our data by sqeqid
        top_n(1, bitscore) %>% # select the top blast search result by bitscore for each seq_id
        dplyr::slice(1) %>%
        ungroup() %>% 
        select(qseqid, category_expanded) %>% 
        rename(top_hit = category_expanded)),
      by = 'qseqid') %>% 
    ggplot(aes(x = qseqid, y = values, fill = groups)) + 
    geom_bar(stat = "identity") +
        viridis::scale_fill_viridis(discrete=T) +
    scale_y_continuous(expand = c(0,0), limits = c(0,10), breaks = c(seq(0, 10, by = 2))) +
    xlab("Psammoneis query sequences (qseqid)") +
    ylab("Top 10 NCBI BlastP hits") +
    labs(fill = "BlastP hit composition") +
    theme(
      axis.text.x = element_blank(),
      axis.text.y = element_text(color = "#000000", size = 10),
      axis.ticks.x = element_blank(),
      # Hide panel borders and remove grid lines
      panel.border = element_blank(),
      # panel.grid.major.x = element_line(colour = "grey", linetype = 3),
      panel.grid.major.y = element_blank(),
      panel.grid.minor = element_blank(),
      # Remove panel background
      panel.background = element_blank(),
      # remove unnecessary text
      axis.title.y = element_markdown(),
      axis.title.x = element_markdown(),
      # Change axis line
      axis.line = element_line(colour = "black"),
      # indicate position of legend
      legend.position = "bottom",
      legend.title = element_text()
    ) +
  facet_wrap(~ top_hit, scales = "free", ncol = 1, labeller = labeller(top_hit = supp.labs))

# let's look at our 17 sequences that made it through NCBI bastP search as bacterial in origin.
blastp_results_final <- 
    left_join(
      (hgt_index_output_orth %>%
         filter(grepl("^Psammoneisjap", qseqid)) %>%
         filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
         filter(bac_genome %in% c("0F","1F","2F","3F")) %>%
         filter(hgt_class == "A")),
      blastp_results_joined,
      by = 'qseqid') %>% 
    group_by(qseqid) %>% 
    top_n(10, bitscore) %>% 
    slice_head(n = 10) %>% 
    ungroup() %>% 
    select(qseqid, category_expanded) %>% 
    table() %>% 
    as.data.frame() %>% 
    pivot_wider(names_from = category_expanded, values_from = Freq) %>% 
    mutate( # now we make a column to calculate the total hits to each qseqid
    total_hits = rowSums(select(., contains(c("Arc","Bac","Euk"))))
    # perc_arc = Archaea/total_hits,
    # perc_bac = Bacteria/total_hits,
    # perc_euk = Eukaryota/total_hits
    ) %>% 
    left_join(
      ., 
      (blastp_results_joined %>% 
        group_by(qseqid) %>% # group our data by sqeqid
        top_n(1, bitscore) %>% # select the top blast search result by bitscore for each seq_id
        dplyr::slice(1) %>%
        ungroup() %>% 
        select(qseqid, category_expanded, sacc, sscinames) %>% 
        rename(top_hit_domain_blastp = category_expanded)),
      by = 'qseqid') %>% 
  rename(archaea_blastp_total_hits = Archaea, 
         bacteria_blastp_total_hits = Bacteria, 
         eukaryota_blastp_total_hits = Eukaryota, 
         total_hits_blastp = total_hits) %>% 
  left_join(., 
            (hgt_index_output_orth_prok_psam %>% 
               select(qseqid, organism_uniprot_d, chrm_bac, bac_gen_peg, bac_gen_gene_annotation, pident_d, bitscore_d, hgt_class, h, h_orth, domain_assignment, orthogroup)),
            by = 'qseqid') %>% 
  filter(top_hit_domain_blastp == "Bacteria")
  
write_csv(blastp_results_final, "~/Downloads/blastp_results_final.csv")

# Investigate those sequences that have a top hit to Eukaryota but also hit to bacteria.
blastp_results_final_composition <- 
  left_join(
    (blastp_results %>% 
       group_by(qseqid) %>% 
       top_n(10, bitscore) %>% 
       slice_head(n = 10) %>% 
       ungroup()), 
    
    (left_join(
               (hgt_index_output_orth %>%
                  filter(grepl("^Psammoneisjap", qseqid)) %>%
                  filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
                  filter(bac_genome %in% c("0F","1F","2F","3F")) %>%
                  filter(hgt_class == "A")),
               blastp_results,
               by = 'qseqid') %>% 
               group_by(qseqid) %>% 
               top_n(10, bitscore) %>% 
               slice_head(n = 10) %>% 
               ungroup() %>% 
               select(qseqid, category_expanded) %>% 
               table() %>% 
               as.data.frame() %>% 
               pivot_wider(names_from = category_expanded, values_from = Freq) %>% 
               mutate( # now we make a column to calculate the total hits to each qseqid
               total_hits = rowSums(select(., contains(c("Arc","Bac","Euk"))))
               ) ),
      by = "qseqid") %>% 
  rename(archaea_blastp_total_hits = Archaea, 
         bacteria_blastp_total_hits = Bacteria, 
         eukaryota_blastp_total_hits = Eukaryota, 
         total_hits_blastp = total_hits) %>% 
  filter(!eukaryota_blastp_total_hits == 10, !bacteria_blastp_total_hits == 10, bacteria_blastp_total_hits >= 5) %>% 
  arrange(qseqid, desc(bitscore))

# now let's add columns to highlight which sequences had only hits to bacterial genomes, those that only the top hit is to our bacterial genomes, and a column to indicate those that did not hit to our bacterial genomes
# You'll need to refer back to your original crisp HGT output file and filter by the IDs that are in the blastp_final DF.

# assign the output directories as variables
donor_directory <-
  "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/02_hgt_index/HGT_index_output/HGT_index_output_donor/"
# list the files in each directory as a variable
donors <-
  list.files(path = donor_directory, full.names = T)
# Here we use fread, lapply, and bind_rows to read in all of our output files and bind them by row (i.e., concatenating them) into a single dataframe for donor and recipient results, respectively.
donor_files_unsliced <-
  hgt_index_output_orth_prok_psam %>% 
  filter(qseqid %in% blastp_results_final$qseqid) %>% 
 # now we will count the number of hits per qseqid,
 # then mutate a new column to show the percentage of hits to our bacterial genomes
  right_join(., bac_gen_unqiue_hits_qseqid, by = c('qseqid', 'sseqid_d', 'orthogroup', 'h', 'h_orth', 'hgt_class', "chrm_bac")) %>% 
  left_join(., 
            (blastp_results_joined %>% 
        group_by(qseqid) %>% # group our data by sqeqid
        top_n(1, bitscore) %>% # select the top blast search result by bitscore for each seq_id
        dplyr::slice(1) %>%
        ungroup() %>% 
        select(qseqid, category_expanded) %>% 
        rename(top_hit_blastp = category_expanded)), 
        by = "qseqid") %>% 
  filter(top_hit_blastp == "Bacteria", !is.na(h)) %>% 
  janitor::remove_empty("cols") %>% 
  select(qseqid, sseqid_d, orthogroup, h, h_orth, domain_assignment, hgt_class, total_hits, genome_pct, top_hit_blastp)
  

```

Finally, we will use NCBI BlastP to check our class A sequences.  
First, we use the SLURM script 'submit_01_download_ncbi_nr_db_c1432.slurm' to download the most recent NR blast database and taxonomy database. 

Next, we export our final 17 class A Psammoneis protein sequences into a FASTA file for protein homology classification using the [PHYRE2](http://www.sbg.bio.ic.ac.uk/phyre2/html/page.cgi?id=index) web service. We will also analyze these proteins using the following services: [TargetP](https://services.healthtech.dtu.dk/service.php?TargetP-2.0) to determine the Subcellular location of our proteins, [SignalP v6.0](https://services.healthtech.dtu.dk/service.php?SignalP) to examine the signal peptides of our sequences, [DeepLOC](https://services.healthtech.dtu.dk/service.php?DeepLoc-2.0) for the prediction of eukaryotic protein subcellular localization using deep learning, [NetGPI](https://services.healthtech.dtu.dk/service.php?NetGPI) for the prediction of GPI-anchoring in our proteins, and [HECTAR](https://webtools.sb-roscoff.fr) to assign proteins to five different categories of subcellular targeting: Signal peptides, type II signal anchors, chloroplast transit peptides, mitochondrion transit peptides and proteins which do not possess any N-terminal target peptide. [ASAFind v.1.1.7](https://bitbucket.org/rocaplab/asafind/src/main/), with [SignalP 3.0](https://services.healthtech.dtu.dk/service.php?SignalP-3.0) short results as input, was used to determine if input proteins are nuclear-encoded plastid proteins in algae with secondary plastids of the red lineage (such as diatoms). 


```{r}
#| label: export Final 17 class A seqs for location protein localization and PHYRE2 analyses
#| eval: true
#| echo: false
#| warning: false
#| error: false
#| output: false

# We will import Psammoneis CDS sequences from orthofinder and filter them to only the sequences of Class A hits to our bacterial genome
psam_CDS <-
  bioseq::read_fasta(file = here("data","02_hgt_index","orthofinder_psammoneis_wade","Psammoneis_japonica.faa"), type = "AA")

# We'll then export those and NCBI blastp them to check that the top hit is prokaryotic
# Psammoneis_only
psam_genic_classA_final <-
  blastp_results_final %>% 
  filter(hgt_class == "A") %>% 
  select(qseqid) %>% 
  tibble()

# now we filter the Psammoneis CDS seqs with the class A qseqids
psam_genic_classA_final_fasta <-
  tibble(label = names(psam_CDS), sequence = psam_CDS) %>% 
  filter(label %in% psam_genic_classA_final$qseqid) %>% 
  mutate(label = paste(">", label, sep = ""))


# split up the psam_CDS_classA_NONbacgen file into chunks so that ncbi BlastP doesn't throw a fit
write.table(psam_genic_classA_final_fasta, 
                 file = here("data","02_hgt_index","orthofinder_psammoneis_wade","Psammoneis_japonica_classA_seqs_final.faa"), 
                          sep = "\n", 
                          quote = FALSE, 
                          row.names = FALSE, 
                          col.names = FALSE
                          )
```

Here we will read in and compile/compare our analyses for PHYRE2 and subcellular localization

```{r}
#| label: import results ffrom subcellular localization and PHYRE2 analyses
#| eval: true
#| echo: false
#| warning: false
#| error: false
#| output: false

phyre2 <- 
  read_delim(here("data","06_subcellular_localization_analyses","phyre2_short_summaryinfo"), delim = " | ") %>% 
  rename("ID" = "Description") %>% 
  janitor::clean_names() %>% 
  arrange(desc(id)) %>% 
  rename_with(.cols = everything(), function(x){paste0("phyre2.", x)}) %>% 
  rename(id = 1) %>% 
  mutate(phyre2.hit_info_1 = str_remove(phyre2.hit_info_1, "^PDB header:"))

signalp <- 
  read_tsv(here("data","06_subcellular_localization_analyses","signalp6_euk_prediction_results.tsv"), comment = "#") %>% 
  janitor::clean_names() %>% 
  arrange(desc(id)) %>% 
  rename_with(.cols = everything(), function(x){paste0("signalp.", x)}) %>% 
  rename(id = 1)

asaFind <- 
  tibble(label = names(psam_CDS), sequence = psam_CDS) %>% 
  filter(label %in% psam_genic_classA_final$qseqid) %>% 
  left_join(., 
            read_tsv(here("data","06_subcellular_localization_analyses","psam_final_asafind_out.txt")),
            by = c("sequence" = "Protein sequence")) %>% 
  select(-Identifier) %>% 
  rename("ID" = "label") %>% 
  janitor::clean_names() %>% 
  arrange(desc(id)) %>% 
  rename_with(.cols = everything(), function(x){paste0("ASAFind.", x)}) %>% 
  rename(id = 1)

targetp <- 
  read_tsv(here("data","06_subcellular_localization_analyses","TargetP_prediction_summary_nonplant.txt"), comment = "#") %>% 
  janitor::clean_names() %>% 
  arrange(desc(id)) %>% 
  rename_with(.cols = everything(), function(x){paste0("targetp.", x)}) %>% 
  rename(id = 1)

netGPI <- 
  read_tsv(here("data","06_subcellular_localization_analyses","NetGPI1.1_output_summary.txt"), comment = "#") %>% 
  janitor::clean_names() %>% 
  arrange(desc(id)) %>% 
  rename_with(.cols = everything(), function(x){paste0("netGPI.", x)}) %>% 
  rename(id = 1)

hectar <- 
  read_tsv(here("data","06_subcellular_localization_analyses","Galaxy2-[HECTAR_results_on_Psammoneis_japonica_classA_seqs_final.faa].tsv"))  %>% 
  dplyr::rename("ID" = "protein id") %>% 
  janitor::clean_names() %>% 
  arrange(desc(id)) %>% 
  rename_with(.cols = everything(), function(x){paste0("heactar.", x)}) %>% 
  rename(id = 1)

deepLOC <- 
  read_csv(here("data","06_subcellular_localization_analyses","DeepLoc2.0_results.csv")) %>% 
  rename("ID" = "Protein_ID") %>% 
  janitor::clean_names() %>% 
  arrange(desc(id)) %>% 
  rename_with(.cols = everything(), function(x){paste0("deepLOC.", x)}) %>% 
  rename(id = 1)

# now let's join all of our DFs into a single one by their ID columns

#put all data frames into list
df_list <- list(phyre2,
                signalp,
                asaFind,
                targetp,
                netGPI,
                hectar,
                deepLOC)

#merge all data frames in list
phyre2_subcell_loc <- 
  df_list %>% 
  reduce(full_join, by = "id") %>% 
   left_join(., 
            (blastp_results_final %>% 
               rename(id = qseqid) %>% 
               select(id, orthogroup)), 
            by = "id") %>% 
  arrange(orthogroup)

write_tsv(
  phyre2_subcell_loc, 
  file = here("data","06_subcellular_localization_analyses","results_combined_final.tsv"))
```

```{r}
#| label: Orthofinder hierarchical orthogroups investigation
#| eval: true
#| echo: false
#| warning: false
#| error: false
#| output: false


PHO_N0 <- 
  read_tsv(here("data","02_hgt_index","orthofinder_psammoneis_wade","OrthoFinder",
                "Results_May12","Phylogenetic_Hierarchical_Orthogroups","N0.tsv")) %>% 
  janitor::clean_names() %>% 
  select(where(~ !(all(is.na(.)) | all(. == "")))) %>% 
  # mutate(across(
  #   .cols = 4:15,
  #   .fns = ~ str_count(.x, ',')+1, 
  #   .names = "{col}_total_genes")) %>% 
  # mutate(across(
  #   .cols = ends_with("_total_genes"), 
  #   .fns = ~ ifelse(is.na(.x), 0, .x))) %>% # go through all of the newly created columns and convert NAs to zeroes
  # mutate(total_genes = rowSums(across(17:27), na.rm = TRUE)) %>% 
  mutate(proteomes_per_OG = apply(across(4:15), 1, count_nonNA_func)) %>% 
  mutate(grouping = 
             case_when(
              is.na(psammoneis_japonica) &
                 is.na(nitzschia_nitz4) &
                 is.na(pseudonitzschia_multistriata) &
                 is.na(fistulifera_solaris) &
                 is.na(pseudonitzschia_multiseries) &
                 is.na(fragilariopsis_cylindrus) &
                 is.na(phaeodactylum_tricornutum) &
                 is.na(psammoneis_japonica) &
                 is.na(seminavis_robusta) &
                 is.na(ectocarpus_siliculosus) &
                 is.na(nannochloropsis_gaditana) ~ "centric",
              is.na(psammoneis_japonica) &
                 is.na(cyclotella_cryptica) &
                 is.na(thalassiosira_pseudonana) &
                 is.na(ectocarpus_siliculosus) &
                 is.na(nannochloropsis_gaditana) ~ "raphid",
               is.na(cyclotella_cryptica) &
                 is.na(thalassiosira_pseudonana) &
                 is.na(ectocarpus_siliculosus) &
                 is.na(nannochloropsis_gaditana) ~ "pennate",
               is.na(ectocarpus_siliculosus) & 
                 is.na(nannochloropsis_gaditana) ~ "diatom"
             )
  ) %>% 
  replace_na(list(grouping = "all")) 

PHO_N4 <- 
  read_tsv(here("data","02_hgt_index","orthofinder_psammoneis_wade","OrthoFinder",
                "Results_May12","Phylogenetic_Hierarchical_Orthogroups","N4.tsv")) %>% 
  janitor::clean_names() %>% 
  select(where(~ !(all(is.na(.)) | all(. == "")))) %>% 
  mutate(raphid_only = ifelse(is.na(psammoneis_japonica),1,0)) %>% 
  add_count(raphid_only)

```

------------------------------------------------------------------------

# bacterial genes of interest and function

For this analysis, we will use the program [METABOLIC](https://github.com/AnantharamanLab/METABOLIC) to predict the metabolic and biogeochemical functional trait profiles of our genomes.

We ran METABOLIC on the AHPCC computing cluster, using default settings and our bacterial genomes as input. The slurm file used for this analysis was 'submit_METABOLIC_768gb.slurm'.

    perl /share/apps/python/anaconda-3.9/envs/METABOLIC_v4.0/METABOLIC/METABOLIC-G.pl -in-gn psam_bac_genomes/ -o metabolic_output/

## Metabolic output exploration and visualization

```{r}
#| label: import METABOLIC output
#| include: false

list.files("/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/03_metabolic_psam_bacteria/metabolic_output", pattern = ".tsv")

metabolic_output <-
  "/Users/cbg/01_projects/psammoneis_genome_and_bacteria/data/03_metabolic_psam_bacteria/metabolic_output/"

metabolic_FunctionHit <-
  read_tsv(paste(metabolic_output,
    "METABOLIC_result_FunctionHit.tsv",
    sep = ""
  ),
  col_names = T
  ) %>%
  # Change present absent to 0 1
  mutate(
    genome_0F = case_when(
    X000000F.quiver_pilon.n_f.Function.presence == "Present" ~ 1,
    X000000F.quiver_pilon.n_f.Function.presence == "Absent" ~ 0),
    genome_1F = case_when(
    X000001F.quiver_pilon.n_f.Function.presence == "Present" ~ 1,
    X000001F.quiver_pilon.n_f.Function.presence == "Absent" ~ 0),
    genome_2F = case_when(
    X000002F.quiver_pilon.n_f.Function.presence == "Present" ~ 1,
    X000002F.quiver_pilon.n_f.Function.presence == "Absent" ~ 0),
    genome_3F = case_when(
    X000003F.quiver_pilon.n_f.Function.presence == "Present" ~ 1,
    X000003F.quiver_pilon.n_f.Function.presence == "Absent" ~ 0),
    ) %>%
  select(-starts_with("x")) %>% 
  mutate(sum = select(., genome_0F:genome_3F) %>% rowSums(na.rm = TRUE)) %>% # this calculates the sum of genomes which possess a metabolic category in their genome.
  filter(!sum == 0) %>% 
  mutate(Total = "")


metabolic_HMMHitNum <-
  read_tsv(paste(metabolic_output,
    "METABOLIC_result_HMMHitNum.tsv",
    sep = ""
  ),
  col_names = T
  ) %>% 
  # Change present absent to 0 1
  mutate(
    genome_0F = case_when(
    X000000F.quiver_pilon.n_f.Hmm.presence == "Present" ~ 1,
    X000000F.quiver_pilon.n_f.Hmm.presence == "Absent" ~ 0),
    genome_1F = case_when(
    X000001F.quiver_pilon.n_f.Hmm.presence == "Present" ~ 1,
    X000001F.quiver_pilon.n_f.Hmm.presence == "Absent" ~ 0),
    genome_2F = case_when(
    X000002F.quiver_pilon.n_f.Hmm.presence == "Present" ~ 1,
    X000002F.quiver_pilon.n_f.Hmm.presence == "Absent" ~ 0),
    genome_3F = case_when(
    X000003F.quiver_pilon.n_f.Hmm.presence == "Present" ~ 1,
    X000003F.quiver_pilon.n_f.Hmm.presence == "Absent" ~ 0),
    ) %>%
  mutate(sum_genome = select(., genome_0F:genome_3F) %>% rowSums(na.rm = TRUE)) %>% # this calculates the sum of genomes which possess a metabolic category in their genome.
  mutate(sum_Hmm = select(., ends_with("Hit.numbers")) %>% rowSums(na.rm = TRUE)) %>% # this calculates the sum of genomes which possess a metabolic category in their genome.
  filter(!sum_genome == 0)
```

# InterProScan results

```{r}
#| label: Import interproscan results
#| eval: true
#| echo: false
#| warning: false
#| error: false
#| output: false

# get the directory our bacterial genome GFF annotation files are in
interpro_directory <-
  here("data","04_interproscan")

# list the gff files in each directory as a variable
interpro_files <-
  list.files(path = interpro_directory, full.names = T, pattern = "\\.tsv")

# Here we use fread, lapply, and bind_rows to read in all of our GFF files and bind them by row (i.e., concatenating them) into a single data frame.
interpro_results <-
  bind_rows(lapply(interpro_files, 
                   fread, 
                   skip=0, 
                   header=FALSE,
                   sep = "\t",
                   fill = TRUE,
                   col.names = c("protein_accession","protein_md5_digest","seq_length","analysis","signature_accession","signature_description","start","stop","score_evalue","status","date","InterPro_annotations_accession","InterPro_annotations_description","GO_annoptations","pathway_annotations"))) %>% 
  filter(protein_accession %in% 
           (hgt_index_output_orth_prok_psam %>% filter(bac_genome %in% c("0F","1F","2F","3F")) %>% filter(hgt_class == "A")%>% tibble())$sseqid_d
         )

# now let's explore what we've got 

```

# _Psammoneis japonica_ and diatom gene family evolution with CAFE
--
##### Software used:
* OrthoFinder (version 2.4.0)
* CAFE (version 5.0.0)
* Python
* treePL

--
##### List of genomes:
| Genome | Version | Proteins |
| ------ | ------- | -------- |
| Cyclotella cryptica | 2.0 | 21,250
| Ectocarpus siliculosus | 1.0 | 16,269
| Fistulifera solaris | 1.0 | 20,426
| Fragilariopsis cylindrus | 1.0 | 21,066
| Nannochloropsis gaditana | 1.0 | 10,929
| Nitzschia sp. Nitz4 | 1.0 | 9,464
| Phaeodactylum tricornutum | 2.0 | 10,408
| Psammoneis japonica | 1.0 | 15,170
| Pseudo-nitzschia multiseries | 1.0 | 19,703
| Pseudo-nitzschia multistriata | 1.0 | 12,152
| Seminavis robusta | 1.0 | 35,996
| Thalassiosira pseudonana | 3.0 | 11,673

### 1. Run OrthoFinder
```
orthofinder -t 12 -a 8 -M msa -S diamond -A mafft -T fasttree -f .
```


### 2. Edit `Orthogroups.GeneCount.tsv` file
* Remove last column titled "Total".
* Add a new first column titled "Desc" that contains gene annotation information or (null).
* Rename species columns to short one-word names.
* Filter the edited file to remove gene families where one or more species have >= 100 gene copies:

```
python3 clade_and_size_filter.py -i Orthogroups.GeneCount.CAFE.txt -o Orthogroups.GeneCount.CAFE.filtered.txt -s
```


### 3. Estimate a species tree
* Use `SpeciesTree_rooted.tree` output by OrthoFinder and STAG.
* Make the species tree ultrametric using treePL.
* Best smoothing parameter is 0.01.

```
treepl treepl_config
```
###### Output: "SpeciesTree_rooted.dated.tree"


### 4. Run CAFE initially under a single lambda model
* Gene families not present in MRCA are filtered: 15503 => 6160.
* Estimate an error model to account for genome assembly or annotation error.

```
cafexp -i Orthogroups.GeneCount.CAFE.filtered.txt -t SpeciesTree_rooted.dated.tree -p -e
```


### 5. Run CAFE with a single lambda and k=2 to k=10 gamma rate categories
* lambda = birth-death rate for gene gain/loss
* alpha = gamma distribution shape parameter
* Poisson distribution for root frequency

```
cafexp -i OrthoGroups.GeneCount.CAFE.filtered.txt -t SpeciesTree_rooted.dated.tre -p -eBase_error_model.txt -o run1 -k 1
```

| Model | -lnL | lambda | alpha |
| ----- | ---------- | ------ | ----- |
| global | 81332.309 | 0.0014 | - |
| k2 | 78053.501 | 0.0023 | 0.61 |
| k3 | 77425.757 | 0.0024 | 0.48 |
| k4 | 77232.543 | 0.0021 | 0.49 |
| k5 | 77157.673 | 0.0019 | 0.51 |
| k6 | 77124.276 | 0.0017 | 0.54 |
| k7 | 77109.302 | 0.0017 | 0.57 |
| **k8** | **77103.864** | **0.0016** | **0.57** |
| k9 | 77107.726 | 0.0016 | 0.65 |
| k10 | 77110.541 | 0.0016 | 0.68 |

##### The global lambda model with k=8 gamma rate categories has the highest -lnL.

##### Do additional runs of the k8 model to ensure convergence


### 6. Run CAFE with two lambda and k=2 to k=10 gamma rate categories
* separate lambda for (1) outgroups, (2) centrics, and (3) pennates

```
cafexp -i OrthoGroups.GeneCount.CAFE.filtered.txt -t SpeciesTree_rooted.dated.tre -eBase_error_model.txt -o multi-lambda -y SpeciesTree_rooted.dated.multi-lambda.tre -p
```

| Model | -lnL | lambda | alpha |
| ----- | ---------- | ------ | ----- |
| multi1 | 80265.478 | 0.0009, 0.0010, 0.0019 | - |
| multi2 | 77090.243 | 0.0012, 0.0017, 0.0034 | 0.67 |
| multi3 | 76857.037 | 0.0009, 0.0011, 0.0022 | 0.50 |

_These multi lambda and multi rate models did not converge between runs._

# Transposable Element density in relation to genic and intergenic HGT candidates

To determine whether our HGT candidates have a higher density of transposable elements than other areas of the genome, I used the the following code and BEDtools. I generated three separte bedfiles for the following: gene (gene annotated), CDS (mRNA annotated), and intergenic ORFs (required additional data processing in R to generate from orfipy header files)
```
# Create a BED file for TE locations. Column 1 = contig name; Column 2 = start position; Column 3 = end position; Column 4 = TE type.
cat psamm_final.round4.all.gff \ # cat the annotation gff file
awk '($3 == "match" && $2 == "repeatmasker") {print $0}' \ # get the lines corresponding to repeat annotations
| awk '/DNA/ || /LTR/ || /LINE/ || /SINE/ ' \ # get the lines containing TE elements
| awk -v OFS='\t' '{ split($9,a,";"); split(a[2],b,":"); split(b[3],c,"%"); print $1,$4,$5,c[1] }' \ # do some text splitting 
> psamm_final.round4.all.TEs.bed

# Create a BED file for gene locations. Column 1 = contig name; Column 2 = start position; Column 3 = end position; Column 4 = gene name.
cat psamm_final.round4.all.gff \ # cat the annotation gff file
| awk '{ if ($3 == "gene") print $0 }' \ # get the lines corresponding to genes
| awk -v OFS='\t' '{ split($9,a,";"); split(a[2],b,"="); print $1,$4,$5,b[2] }' \ # do some text splitting
> psamm_final.round4.all.Genes.bed

# Create a list of contig/scaffold sizes
samtools faidx psam.final_assembly.Quiver_Pilon_Reapr_GapCloser.fa # create samtools index
cut -f 1,2 psam.final_assembly.Quiver_Pilon_Reapr_GapCloser.fa.fai > chrom.sizes # use cut to get the contig names and lengths

# Create a 10kb window around each gene location
bedtools slop \
-i psamm_final.round4.all.Genes.bed \ # the genes BED file
-g chrom.sizes \ # the contig sizes
-b 10000 \ # the size of the flanking window around each gene
> psamm_final.round4.all.Genes.flanking.10kb.bed

# Report the number of overlapping TE elements found in the 10kb window around each gene
bedtools intersect \
-a psamm_final.round4.all.Genes.flanking.10kb.bed \ # the genes BED file with start/end positions that have +10kb
-b psamm_final.round4.all.TEs.bed \ # the TE elements BED file
-c # count the number of TEs overlapping with each gene window
> psamm_final.round4.all.Genes.flanking.10kb.TE.overlapping.tsv # redirect output to a file

```

```{r}
#| label: TE element density
#| eval: true
#| echo: false
#| warning: false
#| error: false
#| output: false

# import general TE density results for all annotated genes
genic_TE_density <- 
  read_tsv(here("data","09_TE_density","psamm_final.round4.all.Genes.flanking.10kb.TE.overlapping.tsv"),
           col_names = c("contig","start_position","end_position","gene_name","TE_density")) %>% 
  mutate(gene_name = paste("Psammoneisjap_",gene_name, sep = ""))

genic_TE_density %>% 
  filter(gene_name %in% blastp_results_final$qseqid) %>% 
  select(TE_density) %>% 
  table()

# Now we import TE desnity for all CDS annotated genes in P. japonica
genic_mRNA_TE_density <- 
  read_tsv(here("data","09_TE_density","psamm_final.round4.all.mRNA.CDS.flanking.10kb.TE.overlapping.tsv"),
           col_names = c("contig","start_position","end_position","gene_name","TE_density")) %>% 
  mutate(gene_name = paste("Psammoneisjap_",gene_name, sep = "")) %>% 
    filter(gene_name %in% blastp_results_final$qseqid) %>% 
  group_by(gene_name) %>% 
    slice_head()

genic_mRNA_TE_density %>% 
  select(TE_density, gene_name) %>% 
  table()


# filter out ORF headers to include only intergenic HGT candidates and then create a BED file of coordinates
intergenic_ORF_BED <-   
  ORF_headers %>%
  filter(qseqid %in% file3_uniprot_prok_hits_0Fthru3F$qseqid) %>% 
  separate(length_orf, into = c("contig_ig_start","contig_ig_stop"), sep = "-", convert = TRUE) %>% 
  separate(coords, into = c("coords_start","coords_stop"), sep = "-", convert = TRUE) %>% 
  mutate(coords_start = coords_start+1 , coords_stop = coords_stop+1) %>% # have to add one to each so that indexing in R is possible. no need to +1 to bedtools range as bedtools uses zero-based indexing. 
  mutate(contig_range = map2(contig_ig_start, contig_ig_stop, `:`)) %>% 
  # now we go through the list with the coords for each ORF
  rowwise() %>% # have to use rowwise to be able to use coords_start/end to pull values out of contig range
  mutate(orf_coord_start = contig_range[coords_start], orf_coord_stop = contig_range[coords_stop]) %>% 
  separate(qseqid, into = c("contig","junk"), sep = ":", remove = FALSE) %>% 
  select(contig, orf_coord_start, orf_coord_stop, qseqid)
  # rowwise %>% # or you can use this to apply the start:end as a list into a new column
  # mutate(contig_range = list(contig_ig_start:contig_ig_stop))

# Write our intergenic ORF bedfile for our ORF HGT candidates
write_tsv(intergenic_ORF_BED, here("data","09_TE_density","psamm_final.round4.all.intergenicORFs.bed"), col_names = FALSE)

# read in the output TE file from our bedtools analyses
intergenic_TE_density <- 
  read_tsv(
    here("data",
         "09_TE_density",
         "psamm_final.round4.all.intergenicORFs.flanking.10kb.TE.overlapping.tsv"),
    col_names = c("contig","start_position","end_position","gene_name","TE_density"))

intergenic_TE_density %>% 
  select(TE_density) %>% 
  table()

```

# Final Figures

To generate the phylogeny for figure 2, we used the species tree inferred by OrthoFinder. The name of their algorithm is STAG (https://doi.org/10.1101/267914) and is described in a preprint. I then dated the species tree using TreePL (https://github.com/blackrim/treePL, https://academic.oup.com/bioinformatics/article/28/20/2689/203074). The input/output files from TreePL are in the zip folder. In the treepl_config.txt file, it lists the commands and you can find listed the min/max calibration times for the MRCA node of the ROOT, DIATOMS, and NITZSCHIA. The min/max times I took from Teofil Nakovs paper for the same corresponding nodes (https://nph.onlinelibrary.wiley.com/doi/full/10.1111/nph.15137).
```{r}
#| label: OG phylogeny with taxonomy labels and repeat percentages
#| eval: true
#| echo: false
#| warning: false
#| error: false
#| output: false

## READ IN DIATOM TREE
tree <- read.tree(here("figs","psammoneis-repeats-figure","SpeciesTree_rooted.dated.tre"))
tree$node.label <- NULL
tree$tip.label <- gsub('_',' ', tree$tip.label)
plot(tree)

## READ IN DATA
table <- read.csv(here("figs","psammoneis-repeats-figure","genome-repeats-summary.csv"),header=T)

## SELECT THE ELEMENT LENGTHS
select <- table %>%
  select(species, coding_bp, repeats_bp, other_bp)

## SELECT THE REPEAT PROPORTIONS
select2 <- table %>%
  select(species, LINE_perc, LTR_perc, DNA_perc,
         Unclass_perc, Simple_perc, other_TE_perc)

## COLORS
colors1 <- c('coding_bp'='black', 'repeats_bp'='grey50', 'other_bp'='grey75')
colors2 <- c('LINE_perc'='#E69F00', 'LTR_perc'='#56B4E9', 'DNA_perc'='#009E73',
             'Simple_perc'='#F0E442', 'other_TE_perc'='#D55E00', 'Unclass_perc'='#CC79A7')

## MAKE BARPLOT OF GENOME CONTENT
p1 <- select %>% pivot_longer(., !c(species), 
                              names_to='Variable',
                              values_to='Value') %>%
  mutate(Variable=factor(Variable)) %>%
  mutate(Variable=fct_relevel(Variable, 'other_bp','repeats_bp','coding_bp')) %>%
  ggplot(aes(x=species, y=Value)) +
  geom_bar(aes(fill=Variable), stat='identity', position='stack') +
  coord_flip() +
  scale_y_continuous(breaks=c(0,50e6,100e6,150e6,200e6), labels=c(0,50,100,150,200)) +
  #scale_x_discrete() +
  scale_fill_manual(values=colors1, labels=c('coding','repetitive','non-coding,\nnon-repetitive')) +
  guides(fill=guide_legend(title='')) +
  theme_minimal() +
  theme(legend.position='right',
        axis.text.y=element_blank(),
        panel.grid.minor.x=element_blank(),
        panel.grid.major.y=element_blank()) +
  #scale_x_discrete(limits=rev) +
  labs(x='', y='Assembly Length (Mb)')


## MAKE BARPLOT OF REPEAT PROPORTIONS
p2 <- select2 %>% pivot_longer(., !c(species), 
                              names_to='Variable',
                              values_to='Value') %>%
  mutate(Variable=factor(Variable)) %>%
  mutate(Variable=fct_relevel(Variable, 'LINE_perc','LTR_perc','DNA_perc',
                              'Simple_perc','other_TE_perc','Unclass_perc')) %>%
  ggplot(aes(y=species, x=Value, fill=Variable)) +
  geom_bar(stat='identity', position='fill') +
  scale_x_continuous() +
  #coord_flip() +
  scale_fill_manual(values=colors2, labels=c('LINE','LTR','DNA',
                                             'simple','others','unclassified')) +
  theme_minimal() +
  theme(axis.text.y=element_blank(),
        panel.grid=element_blank(),
        legend.position='right') +
  guides(fill=guide_legend(title='')) +
  labs(x='Percent of Repetitive', y='')
  #scale_fill_manual(values=colors2) +
  #guides(fill='none') +
  #theme_bw() +
  #theme(legend.position='bottom',
  #      axis.text.y=element_blank()) +
  #labs(x='', y='percent of repetitive elements') +
  #coord_flip()

## PLOT TREE
g <- ggtree(tree, right=T, size=1) +
  coord_cartesian(clip='off') +
  geom_tiplab(size=3.5, align=T, fontface='italic', hjust=-0.05) +
  xlim_tree(115) +
  theme_tree2(plot.margin=margin(0,0,0,0)) +
  xlab('Million Years Ago (Ma)')

g2 <- revts(g) +
  scale_x_continuous(breaks=c(seq(0,-250,-50)), labels=c(seq(0,250,50)))

#g3 <- g2 + geom_text2(aes(subset=!isTip, label=node), hjust=-.3)
g3 <- ggtree::rotate(g2, 17)

g4 <- g3 + geom_label(aes(-120,4.5), label='Raphid\npennates', size=3) +
  geom_label(aes(-65,1), label='Araphid\npennates', size=3) +
  geom_label(aes(-108, 9.5), label='Thalassiosirales', size=3) +
  geom_label(aes(-200, 6.1), label='Diatoms', size=3)

## PLOT TREE AND BARPLOT TOGETHER
#p3 <- p1 %>% insert_left(g4, width=1.5)
#p4 <- p3 %>% insert_right(p2)
p3 <- plot_list(g4, p1, p2, widths=c(0.4,0.35,0.25),
                labels=c('A','B','C'), byrow=T)

p3 <- p2 %>% insert_left(p1, width=2) %>% insert_left(g4, width=3)

ggsave(filename=here("figs","fig_2_OG_phylo_assemblies_repetitives.png"), p3, device='png', height=6, width=12)
```

```{r}
#| label: orf_size_range
#| fig-cap: Figure 8. Density plot of length ranges in nucleotides for intergenic ORFs by HGT classification. Full ORF data for our intergenic HGT analyses can be found in SUPPLEMENTARY FILE X.
#| eval: true
#| warning: false
#| error: false



### ORF size range ###
# maybe add CDS hit ranges?
# file3_uniprot %>%
#   select(length, domain_assignment) %>%
#   mutate(length = as.numeric(length)) %>%
#   ggplot(aes(x = length, color = domain_assignment, label = domain_assignment)) +
#   # geom_density_ridges(rel_min_height = 0.005, alpha = .75) +
#   # geom_density_ridges(alpha=0.75) +
#   viridis::scale_color_viridis(discrete=T) +
#   geomtextpath::geom_textdensity(size = 6, fontface = 2, hjust = 0.2, vjust = 0.3) #
#   scale_x_continuous(limits = c(0,3500), breaks = c(seq(0, 3500, by = 500))) +
#   xlab("ORF length (nucleotides)") +
#   ylab("") +
#   theme(
#     axis.text.x = element_text(angle = 0, hjust = .5, color = "#000000"),
#     axis.text.y = element_text(color = "#000000", size = 10),
#     # Hide panel borders and remove grid lines
#     panel.border = element_blank(),
#     # panel.grid.major.x = element_line(colour = "grey", linetype = 3),
#     panel.grid.major.y = element_blank(),
#     panel.grid.minor = element_blank(),
#     # Remove panel background
#     panel.background = element_blank(),
#     # remove unnecessary text
#     axis.title.y = element_markdown(),
#     axis.title.x = element_markdown(),
#     # Change axis line
#     axis.line = element_line(colour = "black"),
#     # indicate position of legend
#     legend.position = "none",
#     legend.title = element_blank()
#   )
### half-eye dotplot from ggdist
file3_uniprot %>% 
  select(length, domain_assignment) %>% 
  mutate(length = as.numeric(length)) %>% 
  ggplot(aes(y = domain_assignment, x = length, fill = domain_assignment)) +
  stat_slab(scale = 0.7, normalize = "groups") +
  stat_dotsinterval(stroke = 1, point_color = "grey50", interval_color = "grey75", side = "bottom", scale = 0.7, slab_size = NA, layout = "swarm") +
  # geom_density_ridges(alpha=0.75) +
  # viridis::scale_color_viridis(discrete=T) +
  viridis::scale_fill_viridis(discrete=T) +
  scale_x_continuous(limits = c(0,3500), breaks = c(seq(0, 3500, by = 500))) +
  xlab("ORF length (nucleotides)") +
  ylab("") +
  theme(
    axis.text.x = element_text(angle = 0, hjust = .5, color = "#000000"),
    axis.text.y = element_text(color = "#000000", size = 10),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    # panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "none",
    legend.title = element_blank()
  )
```

```{r}
#| label: gIG_domains
#| fig-cap: Figure 7. HGT index classifications for intergenic and genic analyses. Intergenic analyses - A) Frequencies of prokaryotic, indeterminate, and eukaryotic HGT classifications [5] and B) domain level classifications for prokaryotic HGT hits. Genic analyses - C) Frequencies of prokaryotic, indeterminate, and eukaryotic HGT classifications [5], D) domain level classifications for prokaryotic HGT hits, and E) horth classifications [7]. Full datasets for genic and intergenic HGT analyses can be found in SUPPLEMENTARY FILE X and SUPPLEMENTARY FILE X, respectively.
#| eval: true
#| warning: false
#| error: false


### domain classifications for genic and intergenic datasets ###
p1 <-  
  file3_uniprot %>% 
  select(domain_assignment, taxonomic_lineage_superkingdom_d) %>% 
  ggplot(aes(y = domain_assignment, fill = domain_assignment)) +
  geom_bar(width = .4) +
  viridis::scale_color_viridis(discrete=T, begin = 0, end = .95, option = "D") +
  viridis::scale_fill_viridis(name = "HGT index hit classification", discrete=T, begin = 0, end = .95, option = "D") +
  scale_x_continuous(expand = c(0, 0), limits = c(0,1200), breaks = c(seq(0, 1200, by = 300))) +
  xlab("") +
  ylab("") +
  theme(
    axis.text.y = element_text(size = 12, color = "black"),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    # panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    # legend.position = "none",
    # legend.title = element_blank()
  )

p2 <-
  file3_uniprot %>% 
  filter(domain_assignment == "prokaryote" | domain_assignment == "prokaryote_strong") %>% 
  select(domain_assignment, taxonomic_lineage_superkingdom_d) %>% 
  ggplot(aes(y = domain_assignment, fill = taxonomic_lineage_superkingdom_d)) +
  geom_bar(width = .4) +
  scale_fill_manual(name = "Domain of Prokaryote HGT hits", values = c("gray","#696969")) +
  scale_color_manual(values = c("gray","#696969")) +
  scale_x_continuous(expand = c(0, 0), limits = c(0,200), breaks = c(seq(0, 200, by = 50))) +
  xlab("") +
  ylab("") +
  theme(
    axis.text.y = element_text(size = 10, color = "black"),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    # panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    # legend.position = "none",
    # legend.title = element_blank()
  )

p3 <-  
  hgt_index_output_orth %>% 
  ungroup() %>% 
  filter(grepl("^Psammoneisjap", qseqid)) %>% 
  select(domain_assignment, taxonomic_lineage_superkingdom_d) %>% 
  ggplot(aes(y = domain_assignment, fill = domain_assignment)) +
  geom_bar(width = .4) +
  viridis::scale_color_viridis(discrete=T, begin = 0, end = .95, option = "D") +
  viridis::scale_fill_viridis(name = "HGT index hit classification", discrete=T, begin = 0, end = .95, option = "D") +
  scale_x_continuous(expand = c(0, 0)) +
  xlab("Frequency") +
  ylab("") +
  # guides(fill = guide_legend(title = "HGT index hit classification")) +
  theme(
    axis.text.y = element_text(size = 12, color = "black"),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    # panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    # legend.position = "none",
    # legend.title = element_markdown()
  )

p4 <-
  hgt_index_output_orth %>% 
  ungroup() %>% 
  filter(grepl("^Psammoneisjap", qseqid)) %>% 
  filter(domain_assignment == "prokaryote" | domain_assignment == "prokaryote_strong") %>% 
  select(domain_assignment, taxonomic_lineage_superkingdom_d) %>% 
  ggplot(aes(y = domain_assignment, fill = taxonomic_lineage_superkingdom_d)) +
  geom_bar(width = .4) +
  scale_fill_manual(name = "Domain of Prokaryote HGT hits", values = c("gray","#696969")) +
  scale_color_manual(values = c("gray","#696969")) +
  scale_x_continuous(expand = c(0, 0)) +
  xlab("") +
  ylab("") +
  theme(
    axis.text.y = element_text(size = 10, color = "black"),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    # panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    # legend.position = "none",
    # legend.title = element_blank()
  )

p5 <-
  hgt_index_output_orth %>% 
  ungroup() %>% 
  filter(grepl("^Psammoneisjap", qseqid)) %>% 
  filter(domain_assignment == "prokaryote" | domain_assignment == "prokaryote_strong") %>% 
  select(hgt_class, taxonomic_lineage_superkingdom_d) %>% 
  ggplot(aes(y = hgt_class, fill = taxonomic_lineage_superkingdom_d)) +
  geom_bar(width = .4) +
  scale_fill_manual(name = "Domain of Prokaryote HGT hits", values = c("gray","#696969")) +
  scale_color_manual(values = c("gray","#696969")) +
  scale_x_continuous(expand = c(0, 0)) +
  xlab("") +
  ylab("") +
  # guides(fill = guide_legend(title = "Domain of Prokaryote HGT hits")) +
  theme(
    axis.text.y = element_text(size = 10, color = "black"),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    # panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    # legend.position = "none",
    # legend.title = element_markdown()
    )
  

(p1 + inset_element(p2, left = 0.4, bottom = 0.4, right = 1, top = 1, align_to = 'full')) / 
  (p3 + inset_element(p4, left = 0.3, bottom = 0.6, right = .6, top = 1, align_to = 'full') + inset_element(p5, left = 0.6, bottom = 0.6, right = 1, top = 1, align_to = 'full') ) +
  plot_annotation(tag_levels = 'A') +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom", 
    # legend.key.size = unit(0.2, "cm"), 
    legend.title = element_text(size = 11, face = "bold"),
    legend.text = element_text(size = 10)
    )
```

```{r}
#| label: hgt_domains_gIG
#| fig-cap: Figure 6. Ridgeline density plots of h values for genic and intergenic datasets. Colors indicate the respective domain classification for each Diamond BLASTP hit. Vertical gray lines indicate the cutoff h value for determining domain (h = 30). For clarity, plots have been subset to only include values of -200  h  200. Full dataset for genic and intergenic HGT analyses can be found in SUPPLEMENTARY FILE X and SUPPLEMENTARY FILE X, respectively.
#| eval: true
#| warning: false
#| error: false




### HGT values by domain assignment, faceted by genic/intergenic ###
hgt_index_output_orth %>% 
  ungroup() %>% 
  filter(grepl("^Psammoneisjap", qseqid)) %>% 
  select(domain_assignment, h) %>% 
  add_column(g_ig = "genic") %>% 
  rbind(.,
        (file3_uniprot %>% 
          ungroup() %>% 
          select(domain_assignment, h) %>% 
           add_column(g_ig = "intergenic"))) %>% 
  # pivot_longer() +
  ggplot(aes(y = domain_assignment, x = h, color = domain_assignment, fill = domain_assignment)) +
  geom_density_ridges(rel_min_height = 0.005, scale = 2, alpha = 0.7) +
  viridis::scale_color_viridis(discrete=T) +
  viridis::scale_fill_viridis(discrete=T) +
  scale_x_continuous(limits = c(-200,200), breaks = c(0,seq(-200, 200, by = 40))) +
  xlab("HGT index (<i>h</i>)") +
  ylab("") +
  facet_wrap(g_ig~., scales = "free_x")  + # create faceted plots for the g_ig variable
  geom_vline(xintercept = 30, color = "grey", type = 3) +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    # panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "right",
    legend.title = element_blank()
  )
```

```{r}
#| label: blast_hit_gene_annotations
#| fig-cap: Figure 5. Molecular function and gene annotations of genic and intergenic HGT hits, respectively. Genic hits were subset to only include metabolic functions with  5 occurrences. HGT hits that had no annotation are excluded from this figure. Colors indicate the source for the sseqid hit (e.g., yellow for UniProt sequences and all others for individual bacterial genomes). Intergenic hits were subset to only include annotations with > 0 occurrences. Full metabolic annotations for genic and intergenic HGT analyses can be found in SUPPLEMENTARY FILE X and SUPPLEMENTARY FILE X, respectively.
#| eval: true
#| warning: false
#| error: false



### Genic and intergenic metabolic characteristics of potential HGT hits, faceted by genic/intergenic ###
file3_uniprot_prok_hits_eco %>% 
  select(bac_gen_gene_annotation, chrm_bac) %>% 
  # table(useNA="always") %>% # coerce annotations and annotation origins into a tidy count table, while keeping NA's
  table() %>% # coerce annotations and annotation origins into a tidy count table, while keeping NA's
  as.data.frame() %>% 
  dplyr::rename(annotation = bac_gen_gene_annotation, label = chrm_bac) %>% 
  filter(!Freq == 0) %>% 
    
  # rbind our intergenic uniprot only hits  
  rbind(., 
        (
          file3_uniprot_prok_hits_eco %>% 
          select(gene_ontology_molecular_function_uniprot_d) %>%
           mutate(gene_ontology_molecular_function_uniprot_d = str_remove_all(gene_ontology_molecular_function_uniprot_d, " \\[GO\\:[:digit:]+\\]")) %>% 
           table() %>% 
           as.data.frame() %>% 
           # mutate(Var1 = substr(Var1, 1, regexpr(";", Var1, fixed=TRUE)-1)) %>% 
           mutate(annotation = stringi::stri_extract_first_regex(gene_ontology_molecular_function_uniprot_d, "^[^;]+")) %>% 
           add_column(label = "uniprot") %>% 
           select(annotation, label, Freq)
         )) %>% 
  mutate(annotation = forcats::fct_relevel(annotation, sort)) %>% 
  filter(!Freq == 0) %>% 
  add_column(g_ig = "intergenic") %>% 
  
    # rbind our uniprot Genic hits 
  rbind(.,
        (
          hgt_index_output_orth %>% 
          ungroup() %>% 
          filter(grepl("^Psammoneisjap", qseqid)) %>% 
          filter(domain_assignment == "prokaryote"|domain_assignment == "prokaryote_strong") %>% 
          filter(hgt_class %in% c("A","B","C")) %>% 
          select(gene_ontology_molecular_function_uniprot_d) %>% 
          mutate(gene_ontology_molecular_function_uniprot_d = stringi::stri_extract_first_regex(gene_ontology_molecular_function_uniprot_d, "^[^;]+")) %>% 
          table() %>% 
          as.data.frame() %>% 
          rename(Var1 = gene_ontology_molecular_function_uniprot_d) %>% 
          mutate(Var1 = str_remove_all(Var1, " \\[GO\\:[:digit:]+\\]")) %>% 
                   # mutate(Var1 = substr(Var1, 1, regexpr(";", Var1, fixed=TRUE)-1)) %>% 
          dplyr::rename(annotation = Var1) %>% 
          mutate(annotation = forcats::fct_relevel(annotation, sort)) %>% 
          filter(Freq >= 5) %>% 
           add_column(label = "uniprot") %>% 
          add_column(g_ig = "genic")
          )) %>% 
  # rbind our genic bacteria genome hits
  rbind(.,
        (
          hgt_index_output_orth %>% 
          filter(grepl("^Psammoneisjap", qseqid)) %>%
          filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
          filter(hgt_class %in% c("A","B","C")) %>% 
          select(bac_gen_gene_annotation, bac_genome) %>% 
          # table(useNA="always") %>% # coerce annotations and annotation origins into a tidy count table, while keeping NA's
          table() %>% # coerce annotations and annotation origins into a tidy count table, while keeping NA's
          as.data.frame() %>% 
          dplyr::rename(annotation = bac_gen_gene_annotation, label = bac_genome) %>% 
          mutate(annotation = str_remove_all(annotation, " COG3496$|^COG0488\\: |^FIG[:digit:]+\\: | \\(EC [:digit:]+\\.[:digit:]+\\.[:digit:]+\\.+[:digit:]+\\)| \\(EC [:digit:]+\\.[:digit:]+\\.[:digit:]+\\.+\\-\\)")) %>% 
            mutate(annotation = str_replace_all(annotation, 'Hypothetical protein', 'hypothetical protein')) %>% 
            mutate(annotation = str_replace_all(annotation, 'Oxidoreductase', 'oxidoreductase activity')) %>% 
          filter(Freq >= 5) %>% 
          add_column(g_ig = "genic")
        )) %>% 
  group_by(annotation, label, g_ig) %>% 
  summarise_each(funs(sum)) %>% 
  ungroup() %>% 
  group_by(annotation) %>%
  add_count(annotation)%>% 
  ungroup() %>% 
  filter(n >= 1) %>% 
  
  ## Plot it!
  ggplot(aes(y = forcats::fct_rev(annotation), x = Freq, color = label, fill = label)) +
  geom_bar(stat = "identity", width = .6) +
  viridis::scale_color_viridis(discrete=T, begin = 0, end = .95, option = "D") +
  viridis::scale_fill_viridis(discrete=T, begin = 0, end = .95, option = "D", na.value="black") +
  # scale_x_continuous(expand = c(0, 0), limits = c(0,190), breaks = c(190, seq(0, 18, by = 2))) +
  scale_x_continuous(expand = c(0, 0)) +
  # scale_y_discrete(labels = scales::label_wrap(10)) +
  facet_wrap(g_ig~., scales = "free_x")  + # create faceted plots for the g_ig variable
  xlab("Frequency") +
  ylab("") +
  theme(
    axis.text.y = element_text(size = 7.5, color = "black"),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    # panel.grid.major.x = element_line(color = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = c(.95, .12),
    legend.title = element_blank()
  )
```

```{r}
#| label: blast_hit_gene_annotations - bacterial genomes
#| fig-cap: Figure 5. Molecular function and gene annotations of genic and intergenic HGT hits, respectively. Genic hits were subset to only include metabolic functions with  5 occurrences. HGT hits that had no annotation are excluded from this figure. Colors indicate the source for the sseqid hit (e.g., yellow for UniProt sequences and all others for individual bacterial genomes). Intergenic hits were subset to only include annotations with > 0 occurrences. Full metabolic annotations for genic and intergenic HGT analyses can be found in SUPPLEMENTARY FILE X and SUPPLEMENTARY FILE X, respectively.
#| eval: true
#| warning: false
#| error: false



### Genic and intergenic metabolic characteristics of potential HGT hits, faceted by genic/intergenic ###
file3_uniprot_prok_hits_eco %>% 
  select(bac_gen_gene_annotation, chrm_bac) %>% 
  # table(useNA="always") %>% # coerce annotations and annotation origins into a tidy count table, while keeping NA's
  table() %>% # coerce annotations and annotation origins into a tidy count table, while keeping NA's
  as.data.frame() %>% 
  dplyr::rename(annotation = bac_gen_gene_annotation, label = chrm_bac) %>% 
  filter(!Freq == 0) %>% 
  add_column(g_ig = "intergenic") %>% 
    
  # rbind our genic bacteria genome hits
  rbind(.,
        (
          hgt_index_output_orth %>% 
          filter(grepl("^Psammoneisjap", qseqid)) %>%
          filter(domain_assignment == "prokaryote_strong" | domain_assignment == "prokaryote") %>%
          filter(hgt_class %in% c("A","B","C")) %>% 
          select(bac_gen_gene_annotation, bac_genome) %>% 
          # table(useNA="always") %>% # coerce annotations and annotation origins into a tidy count table, while keeping NA's
          table() %>% # coerce annotations and annotation origins into a tidy count table, while keeping NA's
          as.data.frame() %>% 
          dplyr::rename(annotation = bac_gen_gene_annotation, label = bac_genome) %>% 
          mutate(annotation = str_remove_all(annotation, " COG3496$|^COG0488\\: |^FIG[:digit:]+\\: | \\(EC [:digit:]+\\.[:digit:]+\\.[:digit:]+\\.+[:digit:]+\\)| \\(EC [:digit:]+\\.[:digit:]+\\.[:digit:]+\\.+\\-\\)")) %>% 
            mutate(annotation = str_replace_all(annotation, 'Hypothetical protein', 'hypothetical protein')) %>% 
            mutate(annotation = str_replace_all(annotation, 'Oxidoreductase', 'oxidoreductase activity')) %>% 
          filter(Freq >= 1) %>% 
          add_column(g_ig = "genic") 
        )) %>% 
  group_by(annotation, label, g_ig) %>% 
  summarise_each(funs(sum)) %>% 
  ungroup() %>% 
  
  ## Plot it!
  ggplot(aes(y = forcats::fct_rev(annotation), x = Freq, fill = g_ig)) +
  geom_bar(stat = "identity", width = .6) +
  viridis::scale_color_viridis(discrete=T, begin = 0, end = .95, option = "D") +
  viridis::scale_fill_viridis(discrete=T, begin = 0, end = .95, option = "D", na.value="black") +
  # scale_x_continuous(expand = c(0, 0), limits = c(0,190), breaks = c(190, seq(0, 18, by = 2))) +
  scale_x_continuous(labels = scales::number_format(accuracy = 0.1)) +
  # scale_y_discrete(labels = scales::label_wrap(10)) +
  facet_wrap(label~., scales = "free_x", ncol = 4)  + # create faceted plots for the g_ig variable
  xlab("Frequency") +
  ylab("") +
  theme(
    axis.text.y = element_text(size = 7.5, color = "black"),
    axis.text.x = element_text(angle = 45, hjust = 1),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    # panel.grid.major.x = element_line(color = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = c(.95, .2),
    # legend.position = "none",
    legend.title = element_blank()
  )

  # rbind our intergenic uniprot only hits  
file3_uniprot_prok_hits_eco %>% 
          select(gene_ontology_molecular_function_uniprot_d) %>%
           mutate(gene_ontology_molecular_function_uniprot_d = str_remove_all(gene_ontology_molecular_function_uniprot_d, " \\[GO\\:[:digit:]+\\]")) %>% 
           table() %>% 
           as.data.frame() %>% 
           # mutate(Var1 = substr(Var1, 1, regexpr(";", Var1, fixed=TRUE)-1)) %>% 
           mutate(annotation = stringi::stri_extract_first_regex(gene_ontology_molecular_function_uniprot_d, "^[^;]+")) %>% 
           add_column(label = "uniprot") %>% 
           select(annotation, label, Freq) %>% 
  mutate(annotation = forcats::fct_relevel(annotation, sort)) %>% 
  filter(!Freq == 0) %>% 
  add_column(g_ig = "intergenic") %>% 
  
    # rbind our uniprot Genic hits 
  rbind(.,
        (
          hgt_index_output_orth %>% 
          ungroup() %>% 
          filter(grepl("^Psammoneisjap", qseqid)) %>% 
          filter(domain_assignment == "prokaryote"|domain_assignment == "prokaryote_strong") %>% 
          filter(hgt_class %in% c("A","B","C")) %>% 
          select(gene_ontology_molecular_function_uniprot_d) %>% 
          mutate(gene_ontology_molecular_function_uniprot_d = stringi::stri_extract_first_regex(gene_ontology_molecular_function_uniprot_d, "^[^;]+")) %>% 
          table() %>% 
          as.data.frame() %>% 
          rename(Var1 = gene_ontology_molecular_function_uniprot_d) %>% 
          mutate(Var1 = str_remove_all(Var1, " \\[GO\\:[:digit:]+\\]")) %>% 
                   # mutate(Var1 = substr(Var1, 1, regexpr(";", Var1, fixed=TRUE)-1)) %>% 
          dplyr::rename(annotation = Var1) %>% 
          mutate(annotation = forcats::fct_relevel(annotation, sort)) %>% 
          filter(Freq >= 5) %>% 
           add_column(label = "uniprot") %>% 
            add_column(g_ig = "genic")
          )) %>% 
  
  group_by(annotation, label, g_ig) %>% 
  summarise_each(funs(sum)) %>% 
  ungroup() %>% 

  
  ## Plot it!
  ggplot(aes(y = forcats::fct_rev(annotation), x = Freq, color = label, fill = label)) +
  geom_bar(stat = "identity", width = .6) +
  viridis::scale_color_viridis(discrete=T, begin = 0, end = .95, option = "D") +
  viridis::scale_fill_viridis(discrete=T, begin = 0, end = .95, option = "D", na.value="black") +
  # scale_x_continuous(expand = c(0, 0), limits = c(0,190), breaks = c(190, seq(0, 18, by = 2))) +
  scale_x_continuous(expand = c(0, 0)) +
  # scale_y_discrete(labels = scales::label_wrap(10)) +
  facet_wrap(g_ig~., scales = "free_x")  + # create faceted plots for the g_ig variable
  xlab("Frequency") +
  ylab("") +
  theme(
    axis.text.y = element_text(size = 7.5, color = "black"),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    # panel.grid.major.x = element_line(color = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = c(.95, .12),
    legend.title = element_blank()
  )
```

```{r}
#| label: metabolic_plot
#| fig-cap: Figure 9. METABOLIC category (A-C) and function (D-F) output plots for bacterial genomes. A) Absence / Presence of genes in different metabolic categories for each bacterial genome, B) total number of genomes which possess a gene in a metabolic category, and C) total number of genes in each metabolic category for all genomes. D) Absence / Presence of genes in different metabolic functions for each bacterial genome, E) total number of genomes which possess a gene for a metabolic function, and F) total number of genes in each metabolic function for all genomes. Full output data and figures from METABOLIC analysis can be found in SUPPLEMENTARY FILE X and SUPPLEMENTARY FIGURE X, respectively.
#| eval: true
#| warning: false
#| error: false


### METABOLIC Category plots ###
# heat plot of number of genome hits per category
p6 <-
  metabolic_FunctionHit %>%
  filter(!sum == 0) %>%
  melt() %>%
  mutate(Category = replace(Category, Category == "Oxygen metabolism (Oxidative phosphorylation Complex IV)", "Oxygen metabolism")) %>%
  filter(!variable == "sum") %>%
    mutate(across(
      .cols = where(is.numeric),
      factor)) %>% 
  mutate(variable = str_remove(variable, "genome_")) %>% 
  ggplot(aes(variable, fct_rev(Category), fill = value)) +
  geom_tile(aes(fill = value)) +
  scale_fill_manual(name = "Absence / Presence", breaks = c(0,1), values = c("gray","black")) +
  # xlab("Bacterial genomes") +
  ylab("Metabolic category") +
  theme(
    axis.text.x = element_text(angle = 0, hjust = .5, color = "#000000", size = 12),
    axis.text.y = element_text(color = "#000000", size = 12),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_blank(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "bottom",
    legend.title = element_text()
  )

# Tile plot depicting the number of genomes that possess a gene within that category
p7 <-
  metabolic_FunctionHit %>%
  filter(!sum == 0) %>%
      mutate(across(
        .cols = where(is.numeric),
        factor)) %>% 
  ggplot(aes(Total, fct_rev(Category), fill = sum)) +
  geom_tile() +
  scale_fill_manual(name = "Total presence in genomes", 
                    breaks = c(1,2,3,4), 
                    values = c("#440154FF","#31688EFF","#35B779FF","#FDE725FF")) +
  # xlab("Bacterial genomes") +
  ylab("") +
  xlab("") +
  theme(
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.x = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "bottom",
    legend.title = element_text()
  )


# Bar plot depicting the number of genes within that category
p8 <-
  metabolic_FunctionHit %>%
  filter(!sum == 0) %>%
  select(Category) %>%
  table() %>%
  as.data.frame() %>% 
  ggplot(aes(y = fct_rev(Category), x = Freq)) +
  geom_bar(stat = "identity") +
  viridis::scale_fill_viridis(breaks = c(0, 1)) +
  # scale_x_continuous(breaks = c(seq(0, 28, 4))) +
  xlab("No. of genes") +
  ylab("Metabolic category") +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5, color = "#000000"),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "none",
    legend.title = element_blank()
  )


############
### METABOLIC Function plots ###
# heat plot of number of genome hits per feature
p9 <-
  metabolic_FunctionHit %>%
  filter(!sum == 0) %>%
  melt() %>%
  mutate(Category = replace(Category, Category == "Oxygen metabolism (Oxidative phosphorylation Complex IV)", "Oxygen metabolism")) %>%
  filter(!variable == "sum") %>%
    mutate(across(
      .cols = where(is.numeric),
      factor)) %>% 
  mutate(variable = str_remove(variable, "genome_")) %>% 
  ggplot(aes(variable, fct_rev(Function), fill = value)) +
  geom_tile(aes(fill = value)) +
  scale_fill_manual(name = "Absence / Presence", breaks = c(0,1), values = c("gray","black")) +
  # xlab("Bacterial genomes") +
  ylab("Metabolic function") +
  theme(
    axis.text.x = element_text(angle = 0, hjust = .5, color = "#000000", size = 12),
    axis.text.y = element_text(color = "#000000", size = 10),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_markdown(),
    axis.title.x = element_blank(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "bottom",
    legend.title = element_text()
  )

# Tile plot depicting the number of genomes that possess a gene within that category
p10 <-
  metabolic_FunctionHit %>%
  filter(!sum == 0) %>%
      mutate(across(
        .cols = where(is.numeric),
        factor)) %>% 
  ggplot(aes(Total, fct_rev(Function), fill = sum)) +
  geom_tile() +
  scale_fill_manual(name = "Total presence in genomes", 
                    breaks = c(1,2,3,4), 
                    values = c("#440154FF","#31688EFF","#35B779FF","#FDE725FF")) +
  # xlab("Bacterial genomes") +
  ylab("") +
  xlab("") +
  theme(
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.x = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    axis.ticks = element_blank(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "bottom",
    legend.title = element_text()
  )


# Bar plot depicting the number of genes within that category
p11 <-
  metabolic_FunctionHit %>%
  filter(!sum == 0) %>%
  select(-sum) %>% 
  melt() %>% 
  filter(!value == 0) %>% 
  select(Function) %>%
  table() %>%
  as.data.frame() %>% 
  ggplot(aes(y = fct_rev(Function), x = Freq)) +
  geom_bar(stat = "identity") +
  viridis::scale_fill_viridis(breaks = c(0, 1)) +
  # scale_x_continuous(breaks = c(seq(0, 28, 4))) +
  xlab("No. of genes") +
  ylab("Metabolic category") +
  theme(
    axis.text.x = element_text(angle = 0, hjust = 0.5, color = "#000000"),
    # Hide panel borders and remove grid lines
    panel.border = element_blank(),
    panel.grid.major.x = element_line(colour = "grey", linetype = 3),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    # Remove panel background
    panel.background = element_blank(),
    # remove unnecessary text
    axis.title.y = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_markdown(),
    # Change axis line
    axis.line = element_line(colour = "black"),
    # indicate position of legend
    legend.position = "none",
    legend.title = element_blank()
  )

(p6 + p7 + p8) / (p9 + p10 + p11)  +
  plot_annotation(tag_levels = 'A') +
  plot_layout(guides = "collect") &
  theme(
    legend.position = "bottom", 
    # legend.key.size = unit(0.2, "cm"), 
    legend.title = element_text(size = 11, face = "bold"),
    legend.text = element_text(size = 10)
    )
```

## Comparison of metabolic output with genes of interest in bacterial genome .gff files

------------------------------------------------------------------------

# Blobtools figures

Here we use blobtools to generate blob plots from our raw Illumina and PacBio read files.

I used the slurm script '' to perform our analysis and generate the blobplot XXX using the following commands

```

```

# Code appendix

## submit_intergenic_HGT_DMNDblast_uniprot_workflow_psam_768gb.slurm

    #!/bin/bash
    #warning mpi tasks  != slurm tasks 32
    #SBATCH --job-name=ig_dmnd_hgt
    #SBATCH --output=submit_intergenic_HGT_DMNDblast_uniprot_workflow_psam_768gb.slurmOut
    #SBATCH --partition condo
    #SBATCH --constraint aja&0gpu&768gb
    #SBATCH --nodes=1
    #SBATCH --tasks-per-node=24
    #SBATCH --time=3600:00:00

    echo 'cd $SLURM_SUBMIT_DIR'
    cd $SLURM_SUBMIT_DIR

    # here we copy our needed files to a scratch directory with the names of our job ID and then enter that directory to load modules and run our analyses
    echo 'Here we copy our files from our $SLURM_SUBMIT_DIR/ to the scratch directory that contains our job data, /scratch/$SLURM_JOB_ID '
    cp $SLURM_SUBMIT_DIR/psam_bac_genomes/* /scratch/$SLURM_JOB_ID
    cp $SLURM_SUBMIT_DIR/psam_genome/* /scratch/$SLURM_JOB_ID
    cp $SLURM_SUBMIT_DIR/translate_cds.R /scratch/$SLURM_JOB_ID
    cp $SLURM_SUBMIT_DIR/gff_filter_rename_bac.R /scratch/$SLURM_JOB_ID
    cp /home/cbgargas/Diamond_db/*.fasta /scratch/$SLURM_JOB_ID/
    cp -r $SLURM_SUBMIT_DIR/orfipy_psamm_complement_intergenic.fa_out/ /scratch/$SLURM_JOB_ID/

    # now we change our working directory to the scratch job ID folder to work from in there
    echo 'cd /scratch/$SLURM_JOB_ID/'
    cd /scratch/$SLURM_JOB_ID/

    # load modules
    echo 'Load our modules'
    module load bedtools2/2.25.0 gcc/7.3.1 mkl/19.0.4 R/4.0.2 diamond/2.0.1 python/3.8-anaconda

    # source our conda env
    echo 'source our conda env'
    source /share/apps/bin/conda-3.8.sh

    # used this post to figure out the code below
    # https://www.biostars.org/p/216824/

    # use awk to get a bedfile for bedtools that contains only coords for CDS features, then get a fasta of those features for each of our bacterial genomes
    #awk -v OFS='\t' -v FS='\t' '$3 == "CDS" {print $1, $4-1, $5, $1 ":" $4 "-" $5 ":" $9}' 000000F.RAST.n_f.gff \
    #| bedtools getfasta -s -name -fi 000000F.quiver_pilon.n_f.fasta -bed - -fo 000000F_CDS.faa

    # opted to use grep to select lines that contain CDS sequences only
    # made sure to include -s option for bedtools to reverse complement antisense strands
    echo 'get fasta sequences from bacterial genomes using bedtools getfasta'
    Rscript gff_filter_rename_bac.R

    bedtools getfasta -fi 000000F.quiver_pilon.n_f.fasta -bed 0F.gff -fo 000000F.fa -s
    bedtools getfasta -fi 000001F.quiver_pilon.n_f.fasta -bed 1F.gff -fo 000001F.fa -s
    bedtools getfasta -fi 000002F.quiver_pilon.n_f.fasta -bed 2F.gff -fo 000002F.fa -s
    bedtools getfasta -fi 000003F.quiver_pilon.n_f.fasta -bed 3F.gff -fo 000003F.fa -s

    # Check that number of sequences matches manuscript
    echo 'Check that number of bacterial sequences matches manuscript: 15,844'
    grep -c '>' 00000*F.fa

    echo 'check that fasta headers look as wanted.'
    head 00000*F.fa

    echo 'translate bacterial genome sequences'
    Rscript translate_cds.R

    # here we will extract the intergenic regions from our psam genome
    # used the following CLI to drop fasta from end of gff to get result.gff
    # https://www.biostars.org/p/270150/

    #______________________________________

    #Use head with the value we got, minus one, (e.g., line_number - 1) to just give us the gff lines

    head -n 1070595 psamm_final.round4.all.gff > result.gff

    #Sort our .gff file so bedtools doesn't get mad at us.

    bedtools sort -i result.gff > result_sorted.gff

    #Remove contig annotations from our .gff (or else we won't get our complement output)

    grep -v contig result_sorted.gff > result_sorted_final.gff

    #Run bedtools complement to get our intergenic .gff file

    bedtools complement -i result_sorted_final.gff -g sizes_sorted.genome > psamm_complement.txt

    #Run bedtools getfasta on our genome, using the complement .gff file to extract sequences for our intergenic regions.

    bedtools getfasta -fi psam.final_assembly.Quiver_Pilon_Reapr_GapCloser.fa -bed psamm_complement.txt -fo psamm_complement.fa

    # load our orfipy conda env
    echo 'load our orfipy conda env via conda activate orfipy'
    conda activate orfipy

    # Run Orfipy on our intergenic output
    echo 'Run Orfipy on our intergenic output'
    orfipy psamm_complement.fa --pep psamm_complement_ORFs_orfipy.faa --dna psamm_complement_ORFs_orfipy.fa  --strand b --between-stops --include-stop --min 30

    echo 'check number of lines in output .faa file'
    wc -l orfipy_psamm_complement_intergenic.fa_out/psamm_complement_ORFs_orfipy.faa

    echo 'check content of output .faa file for total number of 89F contig ORFs'
    grep -c  89F orfipy_psamm_complement_intergenic.fa_out/psamm_complement_ORFs_orfipy.faa

    echo 'check content of output .faa file for all sequences with 000089F contig ORFs'
    grep -c 000089F orfipy_psamm_complement_intergenic.fa_out/psamm_complement_ORFs_orfipy.faa

    #______________________________________

    #make output directories for output
    mkdir 01_genomes_CDS
    mkdir 02_genomes_CDS_translated

    # copy all output files from R
    mv *.fa 01_genomes_CDS/
    *F.gff* 01_genomes_CDS/
    mv  *.faa 02_genomes_CDS_translated/

    # check contents of each file
    head 01_genomes_CDS/*
    head 02_genomes_CDS_translated/*

    ## Here we construct our Diamond blast database
    # append our CDS AA sequences to our .fna file for blast input
    echo 'append our CDS AA sequences to our .faa file for blast input'
    cat 02_genomes_CDS_translated/00000*F_gff.faa >> uniprot_reviewed_Arc_2157_Bac_2.fasta

    echo 'head nr.faa'
    head 02_genomes_CDS_translated/nr.faa
    echo 'tail nr.faa'
    tail 02_genomes_CDS_translated/nr.faa

    # make diamond db
    echo 'Make Diamond DB using NR + 0-3F bacterial genomes'
    diamond makedb --in uniprot_reviewed_Arc_2157_Bac_2.fasta -d uniprot_prok
    diamond makedb --in uniprot-reviewed_yes+taxonomy_2759_eukaryota.fasta -d uniprot_euk

    mkdir uniprot_dmnd/

    mv *.dmnd uniprot_dmnd/

    ## Here we actually run ourBlastp analysis

    # run Diamond Blastp on intergenic regions
    echo 'run Diamond blastp on intergenic Psammoneis sequences'
    diamond blastp -f 6 -q orfipy_psamm_complement_intergenic.fa_out/psamm_complement_ORFs_orfipy.faa  -d uniprot_dmnd/uniprot_prok.dmnd -k 5 -o intergenic_dmnd_uniprot_prok.out.tsv -v -p 4 -b12 -c1
    diamond blastp -f 6 -q orfipy_psamm_complement_intergenic.fa_out/psamm_complement_ORFs_orfipy.faa  -d uniprot_dmnd/uniprot_euk.dmnd -k 5 -o intergenic_dmnd_uniprot_euk.out.tsv -v -p 4 -b12 -c1

    mkdir dmnd_blastp_hgt_output
    cp *.out.tsv dmnd_blastp_hgt_output

    mkdir DMNDblastp_uniprot_output_psam/

    # copy the output files to directory
    cp -r result.gff result_sorted.gff psamm.genomeFile psamm_complement.txt orfipy_psamm_complement_intergenic.fa_out psamm_complement_edit.bed uniprot_dmnd/ dmnd_blastp_hgt_output/ 02_genomes_CDS_translated/ 01_genomes_CDS/ 03_psam_intergenic/psam_intergenic_and_CDS.faa DMNDblastp_uniprot_output_psam/

    # copy output to submit directory
    cp -r DMNDblastp_uniprot_output_psam/ $SLURM_SUBMIT_DIR/

```{r}
#| label: gff_filter_rename_bac.R
#| eval: false

print("Load our libraries.")
library(tidyverse)

print("Read in our .gff files for 0F-3F.")
gff.0f <-
  read.csv(
    file = "000000F.RAST.n_f.gff",
    header = F,
    sep = "\t", skip = 1
  )

gff.1f <-
  read.csv(
    file = "000001F.RAST.n_f.gff",
    header = F,
    sep = "\t", skip = 1
  )

gff.2f <-
  read.csv(
    file = "000002F.RAST.n_f.gff",
    header = F,
    sep = "\t", skip = 1
  )

gff.3f <-
  read.csv(
    file = "000003F.RAST.n_f.gff",
    header = F,
    sep = "\t", skip = 1
  )

print("Filter our .GFF files to include just CDS annotations & replace contig column to include genome name and peg IDs fr
om column 9.")
gff.0f.filt <-
  gff.0f %>%
  filter(V3 == "CDS")

gff.1f.filt <-
  gff.1f %>%
  filter(V3 == "CDS")

gff.2f.filt <-
  gff.2f %>%
  filter(V3 == "CDS")

gff.3f.filt <-
  gff.3f %>%
  filter(V3 == "CDS")

print("Head(gff.0f.filt)")
head(gff.0f.filt)
print("Head(gff.1f.filt)")
head(gff.1f.filt)
print("Head(gff.2f.filt)")
head(gff.2f.filt)
print("Head(gff.3f.filt)")
head(gff.3f.filt)

print("Check lengths of each .gff file.")
print("Length 0F")
nrow(gff.0f.filt)
print("Length 1F")
nrow(gff.1f.filt)
print("Length 2F")
nrow(gff.2f.filt)
print("Length 3F")
nrow(gff.3f.filt)

print("Check total number of annotations, this value should be equal to 15,844.")
nrow(gff.0f.filt) + nrow(gff.1f.filt) + nrow(gff.2f.filt) + nrow(gff.3f.filt)
15844 == nrow(gff.0f.filt) + nrow(gff.1f.filt) + nrow(gff.2f.filt) + nrow(gff.3f.filt)

print("Finally, we write our filtered dataframes as .GFF files to use with Bedtools getfasta.")
write_delim(gff.0f.filt, file = "0F.gff", delim = "\t", col_names = FALSE)
write_delim(gff.1f.filt, file = "1F.gff", delim = "\t", col_names = FALSE)
write_delim(gff.2f.filt, file = "2F.gff", delim = "\t", col_names = FALSE)
write_delim(gff.3f.filt, file = "3F.gff", delim = "\t", col_names = FALSE)
```

```{r}
#| label: translate_cds.R
#| eval: false

library(tidyverse)

gen0f <-
  bioseq::read_fasta(file = "000000F.fa", type = "DNA")
gen1f <-
  bioseq::read_fasta(file = "000001F.fa", type = "DNA")
gen2f <-
  bioseq::read_fasta(file = "000002F.fa", type = "DNA")
gen3f <-
  bioseq::read_fasta(file = "000003F.fa", type = "DNA")

gen0f
gen1f
gen2f
gen3f

bioseq::seq_nchar(gen0f) %>% range()
bioseq::seq_nchar(gen1f) %>% range()
bioseq::seq_nchar(gen2f) %>% range()
bioseq::seq_nchar(gen3f) %>% range()

gen0faa <-
  bioseq::seq_translate(
    x = gen0f,
    codon_frame = 1,
    code = 1,
    codon_init = FALSE
  )
gen1faa <-
  bioseq::seq_translate(
    x = gen1f,
    codon_frame = 1,
    code = 1,
    codon_init = FALSE
  )
gen2faa <-
  bioseq::seq_translate(
    x = gen2f,
    codon_frame = 1,
    code = 1,
    codon_init = FALSE
  )
gen3faa <-
  bioseq::seq_translate(
    x = gen3f,
    codon_frame = 1,
    code = 1,
    codon_init = FALSE
  )

gen0faa <- tibble(label = names(gen0faa), sequence = gen0faa) %>% mutate(label = paste(">", label, sep = ""))
gen1faa <- tibble(label = names(gen1faa), sequence = gen1faa) %>% mutate(label = paste(">", label, sep = ""))
gen2faa <- tibble(label = names(gen2faa), sequence = gen2faa) %>% mutate(label = paste(">", label, sep = ""))
gen3faa <- tibble(label = names(gen3faa), sequence = gen3faa) %>% mutate(label = paste(">", label, sep = ""))

write.table(gen0faa,
  sep = "\n",
  file = "000000F_gff.faa",
  quote = FALSE, row.names = FALSE, col.names = FALSE
)
write.table(gen1faa,
  sep = "\n",
  file = "000001F_gff.faa",
  quote = FALSE, row.names = FALSE, col.names = FALSE
)
write.table(gen2faa,
  sep = "\n",
  file = "000002F_gff.faa",
  quote = FALSE, row.names = FALSE, col.names = FALSE
)
write.table(gen3faa,
  sep = "\n",
  file = "000003F_gff.faa",
  quote = FALSE, row.names = FALSE, col.names = FALSE
)
```

```{bash diamond_blast_orthogroups_768gb.sh}
#| eval: false

#!/bin/bash

echo 'Run diamond blastp for recipient group'

for f in *.fa
do
  echo "DMND blastp parallelization setup of  $f'!"
  diamond blastp -b12 -c1 --query $f --db uniprot.eukaryota2759.dmnd --out $f.recipient.dmndBlastOut.tsv --outfmt 6 qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore
  mv $f.recipient.dmndBlastOut.tsv HGT_index_output_recipient/
done

echo 'Count the number of files in HGT_index_output_recipient'
ls -halt HGT_index_output_recipient/ | wc -l

echo 'Recipient Loop complete!'

echo 'Run diamond blastp for donor groups'

for f in *.fa
do
  echo "DMND blastp of $f' in parallel!"
  diamond blastp -b12 -c1 --query $f --db uniprot.arc2157.bac2.dmnd --out $f.donor.dmndBlastOut.tsv --outfmt 6 qseqid sseqid pident length mismatch gapopen qstart qend sstart send evalue bitscore
  mv $f.donor.dmndBlastOut.tsv HGT_index_output_donor/
done

echo 'Count the number of files in HGT_index_output_donor'
ls -halt HGT_index_output_donor/ | wc -l

echo 'Donor loop complete! Exiting...'
```

## submit_diamond_HGT_index_768gb.slurm

    #!/bin/bash
    #warning mpi tasks  != slurm tasks 32
    #SBATCH --job-name=hgt_index
    #SBATCH --output=submit_diamond_HGT_index_768gb.slurmOut
    #SBATCH --partition condo
    #SBATCH --constraint aja&0gpu&768gb
    #SBATCH --nodes=1
    #SBATCH --tasks-per-node=24
    #SBATCH --time=3600:00:00
    cd $SLURM_SUBMIT_DIR

    # here we copy our needed files to a scratch directory with the names of our job ID and then enter that directory to load modules and run our analyses
    cp /home/cbgargas/Diamond_db/uniprot_arc_bac_euk_diamond/*.dmnd  /scratch/$SLURM_JOB_ID
    cp -r $SLURM_SUBMIT_DIR/orthofinder_psammoneis_wade/OrthoFinder/Results_May12/Orthogroup_Sequences_Psammoneis/*.fa /scratch/$SLURM_JOB_ID
    cp $SLURM_SUBMIT_DIR/diamond_blast_orthogroups_768gb.sh /scratch/$SLURM_JOB_ID
    cp $SLURM_SUBMIT_DIR/translate_cds.R /scratch/$SLURM_JOB_ID
    cp $SLURM_SUBMIT_DIR/gff_filter_rename_bac.R /scratch/$SLURM_JOB_ID
    cp /home/cbgargas/Diamond_db/*.fasta /scratch/$SLURM_JOB_ID/
    cp -r /home/cbgargas/01_projects/psammoneis_bacteria/psam_bac_genomes/* /scratch/$SLURM_JOB_ID/
    cp -r /home/cbgargas/01_projects/psammoneis_bacteria/psam_genome/* /scratch/$SLURM_JOB_ID/

    # now we change our working directory to the scratch job ID folder to work from in there
    cd /scratch/$SLURM_JOB_ID/

    # load our modules
    echo 'load our modules'
    module load diamond/2.0.1 blast/2.11.0 bedtools2/2.25.0 gcc/7.3.1 mkl/19.0.4 R/4.0.2  gcc impi

    # opted to use grep to select lines that contain CDS sequences only
    # made sure to include -s option for bedtools to reverse complement antisense strands
    echo 'get fasta sequences from bacterial genomes using bedtools getfasta'
    Rscript gff_filter_rename_bac.R

    bedtools getfasta -fi 000000F.quiver_pilon.n_f.fasta -bed 0F.gff -fo 000000F.fa -s
    bedtools getfasta -fi 000001F.quiver_pilon.n_f.fasta -bed 1F.gff -fo 000001F.fa -s
    bedtools getfasta -fi 000002F.quiver_pilon.n_f.fasta -bed 2F.gff -fo 000002F.fa -s
    bedtools getfasta -fi 000003F.quiver_pilon.n_f.fasta -bed 3F.gff -fo 000003F.fa -s

    # Check that number of sequences matches manuscript
    echo 'Check that number of bacterial sequences matches manuscript: 15,844'
    grep -c '>' 00000*F.fa

    echo 'check that fasta headers look as wanted.'
    head 00000*F.fa

    echo 'translate bacterial genome sequences'
    Rscript translate_cds.R

    ## Here we construct our Diamond blast database
    # append our CDS AA sequences to our .fna file for blast input
    echo 'append our CDS AA sequences to our .faa file for blast input'
    cat 00000*F_gff.faa >> uniprot_reviewed_Arc_2157_Bac_2.fasta


    echo 'rm initial bacterial genome files'
    rm 00000*F.fa
    rm psam.final_assembly.Quiver_Pilon_Reapr_GapCloser.fa

    # make diamond db
    echo 'Make Diamond DB using NR + 0-3F bacterial genomes'
    diamond makedb --in uniprot_reviewed_Arc_2157_Bac_2.fasta -d uniprot_prok
    diamond makedb --in uniprot-reviewed_yes+taxonomy_2759_eukaryota.fasta -d uniprot_euk

    mkdir HGT_index_output_donor/
    mkdir HGT_index_output_recipient/
    mkdir HGT_index_output/

    # Start our Diamond blast bash loop
    echo 'Start the diamond blast for loop.'
    ./diamond_blast_orthogroups_768gb.sh

    cd HGT_index_output_donor/
    for f in *.tsv; do sed -i "s/$/\t$f/" $f; done
    for f in *.tsv; do sed -i '1 s/^/qseqid\tsseqid_d\tpident_d\tlength_d\tmismatch_d\tgapopen_d\tqstart_d\tqend_d\tsstart_d\tsend_d\tevalue_d\tbitscore_d\torthogroup\n/' $f; done

    cd ../HGT_index_output_recipient/
    for f in *.tsv; do sed -i "s/$/\t$f/" $f; done
    for f in *.tsv; do sed -i '1 s/^/qseqid\tsseqid_r\tpident_r\tlength_r\tmismatch_r\tgapopen_r\tqstart_r\tqend_r\tsstart_r\tsend_r\tevalue_r\tbitscore_r\torthogroup\n/' $f; done

    cd ..

    echo 'Move recipient and donor directories to HGT_index_output'
    mv HGT_index_output_* HGT_index_output/

    # copy output to submit directory
    echo 'Copy Output directories to submit directory'
    cp -r HGT_index_output $SLURM_SUBMIT_DIR/

## submit_METABOLIC_768gb.slurmOut

    # srun diamond


    #!/bin/bash
    #warning mpi tasks  != slurm tasks 32
    #SBATCH --job-name=metabolic
    #SBATCH --output=submit_METABOLIC_768gb.slurmOut
    #SBATCH --partition condo
    #SBATCH --constraint aja&0gpu&192gb
    #SBATCH --nodes=1
    #SBATCH --tasks-per-node=8
    #SBATCH --time=3600:00:00

    cd $SLURM_SUBMIT_DIR

    # here we copy our needed files to a scratch directory with the names of our job ID and then enter that directory to load modules and run our analyses
    cp -r /home/cbgargas/01_projects/psammoneis_bacteria/psam_bac_genomes/ /scratch/$SLURM_JOB_ID/

    # now we change our working directory to the scratch job ID folder to work from in there
    cd /scratch/$SLURM_JOB_ID/

    # load modules and conda environments
    module load python/anaconda-3.9
    source /share/apps/bin/conda-3.9.sh
    conda activate METABOLIC_v4.0

    perl /share/apps/python/anaconda-3.9/envs/METABOLIC_v4.0/METABOLIC/METABOLIC-G.pl -in-gn psam_bac_genomes/ -o metabolic_output/

    # copy output to submit directory
    cp -r metabolic_output/ $SLURM_SUBMIT_DIR/
